% !TEX root = ../notes.tex

\documentclass[../notes.tex]{subfiles}

\begin{document}

We began class with some remarks on $L^\infty(X)$, which I have just edited into last class's notes.

\subsection{The Moral}
Let $X$ be a compact Hausdorff topological space $X$. We refer to \cite{elber-top} and in particular subsection~4.2.1 for some facts about compact Hausdorff spaces.
\begin{notation}
	Fix a compact Hausdorff topological space $X$. Then $C(X)$ is the $\CC$-vector space of continuous functions $X\to\CC$.
\end{notation}
\begin{remark}
	We note $C(X)$ can be upgraded to a normed vector space with the norm
	\[\norm f_{C(X)}\coloneqq\sup_{x\in X}\left|f(x)\right|.\]
	We will omit the checks that $\norm\cdot_{C(X)}$ is actually a norm, though we note that it is basically $\norm\cdot_\infty$. Because $C(X)$ can be checked to be complete with respect this norm (convergence for $\norm\cdot_{C(X)}$ means uniform convergence, which preserves continuity), we see that $C(X)$ is a Banach space.
\end{remark}
The Riesz representation theorem is basically the statement that $C(X)^*$ is exactly the space of complex Borel measures on $X$. (Notably, complex Borel measures on $X$ form a $\CC$-vector space.) This claim is not literally true, but we will be able to make it true enough.

\subsection{Locally Compact Spaces}
It will be helpful to have some understanding of locally compact spaces. We borrow the following exposition from \cite[subsection~4.8.1]{elber-top}.
\begin{definition}[Locally compact]
	A topological space $(X,\mc T)$ is \textit{locally compact} if and only if each point $x\in X$ has some open subset $U\in\mc T$ containing $x$ such that $\overline U$ is compact.
\end{definition}
\begin{example}
	The set of real numbers $\RR$ with the usual topology is locally compact. Indeed, any $x\in\RR$ has the open neighborhood $(x-1,x+1)$ with closure $[x-1,x+1]$, and $[x-1,x+1]$ is compact.
\end{example}
\begin{example}
	For the same reason, the space $[a,b)$ is also locally compact.
\end{example}
\begin{example}
	Finite-dimensional Euclidean spaces $V$ are locally compact spaces; for example, they can be embedded into the compact projective space $\PP V$. However, infinite-dimensional Banach spaces fail to be locally compact. Indeed, up to shrinking, this would imply that sufficiently small open balls are compact, which is false.
\end{example}
Here are some helpful facts.
\begin{lemma} \label{lem:better-loc-compact}
	Fix a locally compact Hausdorff space $(X,\mc T)$. Then any $x\in X$ and open subset $U\in\mc T$ containing $x$ has some open subset $U_x\subseteq X$ containing $x$ such that $\overline{U_x}$ is compact and $\overline{U_x}\subseteq U$.
\end{lemma}
\begin{proof}
	We begin by finding our promised $U'$ containing $x$ with $\overline{U'}$ compact. Thus, it suffices to find some open subset $V$ containing $x$ such that $\overline V$ is compact and $\overline V\subseteq U\cap U'$, but now we see that
	\[\overline{U\cap U'}\subseteq\overline{U'}\]
	is a closed subset of the compact space $\overline{U'}$ and therefore compact. In particular, we can replace $U$ with $U\cap U'$ and assume that $\overline U$ is compact.

	Now, let $\del U\coloneqq\overline U\setminus U$ be the boundary of $U$. Notably, $\del U$ is a closed subset of the compact space $\overline U$, so $\del U$ is compact. Because $\{x\}$ is a closed subset in $U$ (note $X\setminus\{x\}$ is open, so $U\setminus\{x\}$ is open in the relative topology), the fact that compact Hausdorff spaces are normal grants open subsets $U_x$ and $U_{\del}$ of $\overline U$ with $x\in U_x$ and $\del U\subseteq U_\del$.

	Now, $U_x\subseteq\overline U\setminus U_\del\subseteq\overline U\setminus{\del U}$, so we see $\overline{U_x}\subseteq\overline U\setminus U_\del$ because $\overline U\setminus U_\del$ is a closed subset of $\overline U$. Further, $\overline{U_x}$ is a closed subset of a compact space $\overline U$, so $\overline{U_x}$ is compact, so we are done.
\end{proof}
\begin{remark}
	\Cref{lem:better-loc-compact} basically says that open subspaces of locally compact Hausdorff spaces are locally compact.
\end{remark}
We can extend the previous result past points to full compact sets.
\begin{lemma} \label{lem:best-loc-compact}
	Fix a locally compact Hausdorff space $(X,\mc T)$ and some compact subset $C\subseteq X$. Then any open subset $U$ containing $C$ has some open subset $U_C$ containing $C$ such that $\overline{U_C}$ is compact and $\overline{U_C}\subseteq U$.
\end{lemma}
\begin{proof}
	We use \Cref{lem:better-loc-compact}. For each $x\in C$, find some $U_x$ by \Cref{lem:better-loc-compact} with $U_x$ containing $x$ with $\overline{U_x}$ compact and $\overline{U_x}\subseteq U$. Then we see that
	\[C\subseteq\bigcup_{x\in C}U_x,\]
	so we have provided an open cover for $C$, so we can choose finitely many $\{x_i\}_{i=1}^n\subseteq C$ with $U_i\coloneqq U_{x_i}$ so that
	\[C\subseteq\bigcup_{i=1}^nU_i\subseteq U_C.\]
	Now, we see that
	\[\overline{\bigcup_{i=1}^nU_i}=\bigcup_{i=1}^n\overline{U_i}\]
	is a compact subset of $U$ because being compact is closed under finite unions, so $\bigcup_{i=1}^nU_i$ is the required open subset.
\end{proof}

\end{document}