% !TEX root = ../notes.tex

\documentclass[../notes.tex]{subfiles}

\begin{document}

\section{Local Fields}

In this section, we introduce the main characters of our intermission, which are the characteristic-$0$ local fields.

\subsection{The \texorpdfstring{$p$}{ p}-adics}
Fix a prime $p$ in the following discussion. The goal of the present subsection is to define the ring of $p$-adic integers $\ZZ_p$. Approximately speaking, the point will be that $\ZZ_p$ is able to capture the information of $\ZZ/p\ZZ$ and $\ZZ/p^2\ZZ$ and $\ZZ/p^3\ZZ$ and so on, all at once.

To see why this might be helpful, we recall the idea of ``local obstructions'' we saw much earlier in this course: an equation might have solutions $\ZZ/p\ZZ$ but lose solutions in $\ZZ/p^2\ZZ$, or it might have solutions in $\ZZ/p^3\ZZ$ but lose solutions in $\ZZ/p^4\ZZ$.
\begin{example}
	The equation $x^2+1=0$ has solutions in $\ZZ/2\ZZ$ but no solutions in $\ZZ/4\ZZ$.
\end{example}
\begin{example}
	The equation $4x^2+4=0$ has solutions in $\ZZ/8\ZZ$ but no solutions in $\ZZ/16\ZZ$.
\end{example}
The ring $\ZZ_p$ will be able to keep track of all this modular information at once. For example, thinking about $\ZZ_p$ will allow us to cleanly thing about statements such as the following one.
\begin{restatable}{proposition}{ztwosquares} \label{prop:squares-z2}
	Let $a$ be an odd integer. The following are equivalent.
	\begin{listalph}
		\item The equation $x^2\equiv a\pmod 8$ has solutions.
		\item The equation $x^2\equiv a\pmod{2^\nu}$ for any power $2^\nu$.
	\end{listalph}
\end{restatable}
\noindent With this in mind, a good first guess for $\ZZ_p$ would be the infinite product ring
\[R_p\coloneqq\prod_{\nu=0}^\infty\ZZ/p^\nu\ZZ.\]
However, this ring has many problems. For example, $R_p$ is not an integral domain: $(1,0,0,\ldots)\cdot(0,1,0,\ldots)=(0,0,0,\ldots)$. A worse problem is that it somehow fails to actually care about modular information: a central property of the $\ZZ/p^\nu\ZZ$s is that if an equation has solutions in $\ZZ/p^{\nu+1}\ZZ$, then it will have solutions in $\ZZ/p^\nu\ZZ$ because of the reduction map
\begin{equation}
	\ZZ/p^{\nu+1}\ZZ\to\ZZ/p^\nu\ZZ. \label{eq:zpz-proj}
\end{equation}
This reduction map would allow us to recover $\ZZ/p^\nu\ZZ$-solutions from ``higher-order'' solutions, and we would like $\ZZ_p$ to mirror this structure.

In particular, we would like projection maps $\ZZ_p\to\ZZ/p^\nu\ZZ$ for each $p^\nu$, and we want these maps to commute with the maps \eqref{eq:zpz-proj}. This allows us to fix $R_p$ into the following ring.
\begin{defihelper}[$p$-adic integers] \nirindex{p-adics integers@$p$-adic integers}
	Fix a prime $p$. Then we define the ring of \textit{$p$-adic integers} by
	\[\ZZ_p\coloneqq\Bigg\{(a_\nu)_{\nu=0}^\infty\in\prod_{\nu=0}^\infty\ZZ/p^\nu\ZZ:a_{\nu+1}\equiv a_\nu\pmod{p^\nu}\Bigg\}.\]
	In practice, it is helpful to view $(a_\nu)_\nu$ as a sequence of integers, but it is important to remember that we only care about $a_\nu\pmod{p^\nu}$.
\end{defihelper}
\begin{remark} \label{rem:upgrade-zp-equiv}
	Quickly, note that $a_{\nu+1}\equiv a_\nu\pmod{p^\nu}$ implies (by induction) that
	\[a_{\nu+N}\equiv a_\nu\pmod{p^\nu}\]
	for any nonnegative integer $N\ge0$. Indeed, there is nothing to say for $N=0$, and then the inductive step recognizes that $a_{\nu+N+1}\equiv a_{\nu+N}\pmod{p^{\nu+N}}$ and hence $a_{\nu+N+1}\equiv a_{\nu+N}\equiv a_\nu\pmod{p^\nu}$.
\end{remark}
\begin{example} \label{ex:power-series-z2}
	Fix $p=2$ and define $a_\nu\coloneqq1+2+2^2+\cdots+2^{\nu-1}$. Then $(a_\nu)_\nu\in\ZZ_p$.
\end{example}
\begin{proof}
	Certainly we have defined a sequence of elements $a_\nu\in\ZZ/2^\nu\ZZ$. It remains to check that
	\[a_{\nu+1}=\sum_{k=0}^\nu2^k\equiv\sum_{k=0}^{\nu-1}2^k=a_\nu\pmod{2^\nu},\]
	so we are done.
\end{proof}
Of course, we can write elements of $\ZZ_p$ as sequences in $\prod_\nu\ZZ/p^\nu\ZZ$, but \Cref{ex:power-series-z2} suggests a more intuitive way to look at our elements: any element of $\ZZ_p$ can be thought of as a power series
\[c_0+c_1p+c_2p^2+\cdots\]
where $c_0,c_1,c_,\ldots\in\{0,1,\ldots,p-1\}$. Let's explain this bijection.
\begin{proposition} \label{prop:p-adic-power-series}
	Fix a prime $p$. Then there is a unique bijection $\{0,1,\ldots,p-1\}^\NN\to\ZZ_p$ given by sending $(c_\nu)_{\nu=0}^\infty$ to the sequence
	\[\Bigg(\sum_{k=0}^{\nu-1}c_kp^k\Bigg)_{\nu=0}^\infty\in\ZZ_p.\]
\end{proposition}
\begin{proof}
	Let the function $\{0,1,\ldots,p-1\}^\NN\to\ZZ_p$ be called $\varphi$. Here are our checks.
	\begin{itemize}
		\item Well-defined: given $(c_\nu)_{\nu=0}^\infty$, we must show that $\varphi\left((c_\nu)_{\nu=0}^\infty\right)\in\ZZ_p$. Well, expanding out, it is enough to note that
		\[\sum_{k=0}^{\nu}c_kp^k\equiv\sum_{k=0}^{\nu-1}c_kp^k\pmod{p^\nu},\]
		completing the argument.
		\item Injective: suppose we have two sequences $(c_\nu)_{\nu=0}^\infty$ and $(d_\nu)_{\nu=0}^\infty$ going to the same element in $\ZZ_p$. Then checking in $\ZZ_p$ tells us that
		\[\sum_{k=0}^{\nu-1}c_kp^k\equiv\sum_{k=0}^{\nu-1}d_kp^k\pmod{p^\nu}\]
		for any $\nu$. We now claim that $c_\nu=d_\nu$ for each $\nu\ge0$ by induction on $k$. If $\nu=0$, then we note $c_0\equiv d_0\pmod p$ enforces $c_0=d_0$ because $c_0,d_0\in\{0,1,\ldots,p-1\}$. Now, for the inductive step, suppose $c_k=d_k$ for $k<\nu$, then we see that
		\[\sum_{k=0}^{\nu}c_kp^k\equiv\sum_{k=0}^{\nu}d_kp^k\pmod{p^{\nu+1}},\]
		but we already know that
		\[\sum_{k=0}^{\nu-1}c_kp^k=\sum_{k=0}^{\nu-1}d_kp^k\]
		by the inductive hypothesis, so we conclude $c_\nu p^\nu\equiv d_\nu p^\nu\pmod{p^{\nu+1}}$, so $c_\nu\equiv d_\nu\pmod p$, so $c_\nu=d_\nu$.
		\item Surjective: fix any $(a_\nu)_{\nu=0}^\infty\in\ZZ_p$. Now, we would like to have
		\begin{equation}
			a_\nu\stackrel?\equiv\sum_{k=0}^{\nu-1}c_kp^k\pmod{p^\nu} \label{eq:desired-power-series}
		\end{equation}
		for each $\nu\ge0$. We will define $(c_\nu)_{\nu=0}^\infty$ inductively to satisfy these. We begin by choosing $c_0\equiv a_1\pmod p$, which is possible because we may choose $c_0\in\{0,1,\ldots,p-1\}$.

		For the inductive step, suppose we are given $c_0,\ldots,c_{\nu-1}$ so that \eqref{eq:desired-power-series} holds. Observe that this also implies
		\[a_{\nu+1}\equiv\sum_{k=0}^{\nu-1}c_kp^k\pmod{p^\nu},\]
		so we define $c_\nu\in\{0,1,\ldots,p-1\}$ by
		\[c_\nu\equiv\frac1{p^\nu}\Bigg(a_{\nu+1}-\sum_{k=0}^{\nu-1}c_kp^k\Bigg)\pmod p.\]
		Rearranging yields
		\[a_{\nu+1}\equiv\sum_{k=0}^\nu c_kp^k\pmod{p^{\nu+1}},\]
		completing the inductive step.
		\qedhere
	\end{itemize}
\end{proof}

\subsection{Ring Theory of \texorpdfstring{$\ZZ_p$}{ Zp}}
Here are some basic checks on $\ZZ_p$.
\begin{lemma}
	Fix a prime $p$. Then $\ZZ_p$ is a subring of $R_p$.
\end{lemma}
\begin{proof}
	We have the following checks.
	\begin{itemize}
		\item Identities: note that $(0,0,\ldots)\in\ZZ_p$ and $(1,1,\ldots)\in\ZZ_p$ because $a\equiv a\pmod{p^\nu}$ for any $p^\nu$ and $a\in\{0,1\}$.
		\item Closure: given $(a_\nu)_\nu,(b_\nu)_\nu\in\ZZ_p$, we note that
		\begin{align*}
			a_{\nu+1}+b_{\nu+1} &\equiv a_\nu+b_\nu\pmod{p^\nu}, \\
			a_{\nu+1}\cdot b_{\nu+1} &\equiv a_\nu\cdot b_\nu\pmod{p^\nu},
		\end{align*}
		so $(a_\nu+b_\nu)_\nu,(a_\nu+b_\nu)\in\ZZ_p$ as well.
		\qedhere
	\end{itemize}
\end{proof}
We complained that $R_p$ is not an integral domain, but $\ZZ_p$ is.
\begin{lemma} \label{lem:zp-int-domain}
	Fix a prime $p$. Then $\ZZ_p$ is an integral domain.
\end{lemma}
\begin{proof}
	The point is to use the coherence of the coordinates of in elements of $\ZZ_p$. Suppose we have nonzero $(a_\nu)_\nu,(b_\nu)_\nu\in\ZZ_p$, and we show that the product is nonzero. Because $(a_\nu)_\nu$ and $(b_\nu)_\nu$ are nonzero, we can find some specific nonnegative integers $m$ and $n$ such that $a_m\not\equiv0\pmod{p^m}$ and $b_n\not\equiv0\pmod{p^n}$.
	
	Thus, the largest power of $p$ dividing $a_m$ is less than $m$, and the same then holds for $a_{m+n}$; similarly, and the largest power of $p$ dividing $b_n$ is less than $n$, and the same then holds for $b_{m+n}$. Thus, the largest power of $p$ dividing $a_{m+n}b_{m+n}$ is less than $m+n$, so
	\[a_{m+n}b_{m+n}\not\equiv0\pmod{p^{m+n}},\]
	so $(a_\nu)_\nu\cdot(b_\nu)_\nu\ne0$.
\end{proof}
It is notable that there is a copy of $\ZZ$ in $\ZZ_p$, and $\ZZ$ is ``dense'' in $\ZZ_p$ in some sense. This tells us that $\ZZ_p$ is indeed a fairly good approximation of $\ZZ$, but do remember that $\ZZ_p$ only cares about$\pmod{p^\nu}$ information.
\begin{lemma}
	Fix a prime $p$. For any $(a_\nu)_\nu\in\ZZ_p$ and positive integer $N$, there exists an integer $a\in\ZZ$ such that
	\[a\equiv a_\nu\pmod{p^\nu}\]
	for all $\nu\le N$.
\end{lemma}
\begin{proof}
	Choose $a\coloneqq a_N$. Then any $\nu\le N$ has
	\[a_N\equiv a_\nu\pmod{p^\nu}\]
	by construction of $\ZZ_p$, so we are done.
\end{proof}
\begin{lemma} \label{lem:z-into-zp}
	Fix a prime $p$. The map $i_p\colon\ZZ\to\ZZ_p$ given by $i_p(a)\coloneqq(a)_{\nu=0}^\infty$ is an injective ring homomorphism.
\end{lemma}
\begin{proof}
	Note that the maps $\ZZ\to\ZZ/p^\nu\ZZ$ for each $p^\nu$ are all (surjective) ring homomorphisms, so we have a ring map
	\[\ZZ\to\prod_{\nu=0}^\infty\ZZ/p^\nu\ZZ\]
	given by $a\mapsto(a)_{\nu=0}^\infty$. So to check that $i_p$ is a ring homomorphism, we merely need to check that the above map lands in $\ZZ_p$. Well, we see that
	\[a\equiv a\pmod{p^\nu}\]
	for any $p^\nu$, so indeed $i_p(a)\in\ZZ_p$.

	Lastly, we check that $i_p$ is injective. It suffices to check that $i_p$ has trivial kernel. Well, suppose we have some nonzero integer $n\in\ZZ$. By unique prime factorization in $\ZZ$, there is a some largest nonnegative integer $\nu$ such that $p^\nu\mid n$. But then
	\[n\not\equiv0\pmod{p^{\nu+1}},\]
	so it follows that $i_p(n)\ne0$.
\end{proof}
\begin{remark}
	\Cref{lem:z-into-zp} is somewhat remarkable in that each individual map $\ZZ\to\ZZ/p^\nu\ZZ$ has lost information (it has nontrivial kernel), but gluing them all together does not lose any information about $\ZZ$! Typically, we think about taking modding out as forgetting some information, but we can technically recover $\ZZ$ from within $\ZZ_p$.
\end{remark}
We now make a few remarks on the ring structure of $\ZZ_p$, simply because it is so nice. To emphasize that $\ZZ_p$ only cares about the prime $p$, we note that all other primes are invertible in $\ZZ_p$. In fact, we have the following stronger statement.
\begin{proposition} \label{prop:units-in-zp}
	Fix a prime $p$ and some $(a_\nu)_\nu\in\ZZ_p$. The following are equivalent.
	\begin{listalph}
		\item $(a_\nu)_\nu$ is a unit.
		\item $a_\nu\pmod{p^\nu}$ is a unit for all $\nu\ge1$.
		\item $a_\nu\pmod{p^\nu}$ is a unit for some $\nu\ge1$.
	\end{listalph}
\end{proposition}
\begin{proof}
	We show our implications in sequence.
	\begin{itemize}
		\item We show (a) implies (b). Suppose that $(a_\nu)_\nu$ is a unit so that we have some $(b_\nu)_\nu\in\ZZ_p$ such that $(a_\nu)_\nu\cdot(b_\nu)_\nu=1$. Here, $1$ is the sequence $(1,1,\ldots)$, so we see that
		\[a_\nu\cdot b_\nu\equiv1\pmod{p^\nu}\]
		for each $\nu$, which is (b).
		\item Note that (b) implies (c) with no work.
		\item We show (c) implies (b). Suppose that $a_N\pmod{p^N}$ is a unit where $N\ge1$. Now, for any $\nu\ge1$, we see that
		\[a_N\equiv a_\nu\pmod{p^{\min\{N,\nu\}}},\]
		so $a_N\equiv a_\nu\pmod p$, so $p\nmid a_N$ implies that $p\nmid a_\nu$. But $p\nmid a_\nu$ implies that $a_\nu$ is a unit in $(\ZZ/p^\nu\ZZ)^\times$, which is what we wanted.
		\item We show that (b) implies (a). By hypothesis, we may define
		\[b_\nu\equiv\begin{cases}
			1\pmod1 & \text{if }\nu=0, \\
			a_\nu^{-1}\pmod{p^\nu} & \text{if }\nu\ge1.
		\end{cases}\]
		Quickly, note that $(a_\nu)_\nu\cdot(b_\nu)_\nu=1$, so it only remains to verify that $(b_\nu)_\nu$ lives in $\ZZ_p$. Well, for any $n\ge m$, we note that
		\[1\equiv b_na_n\equiv b_na_m\pmod{p^m},\]
		so $b_n\equiv a_m^{-1}\equiv b_m\pmod{p^m}$, as needed.
		\qedhere
	\end{itemize}
\end{proof}
\begin{lemma} \label{lem:principal-ideals-zp}
	Fix a prime $p$. For any integer $n\ge1$, we have
	\[p^n\ZZ_p=\left\{(a_\nu)_\nu\in\ZZ_p:a_n\equiv0\pmod{p^n}\right\}.\]
\end{lemma}
\begin{proof}
	There are two inclusions to show. In one direction, for any $(a_\nu)_\nu\in\ZZ_p$, we see that $p^n(a_\nu)_\nu=\left(p^na_\nu\right)_\nu$ has $p^na_\nu\equiv0\pmod{p^n}$. In the other direction, suppose that $(a_\nu)_\nu\in\ZZ_p$ has $a_n\equiv0\pmod{p^n}$. Then we use \Cref{prop:p-adic-power-series} to expand $(a_\nu)_\nu$ as a power series
	\[\sum_{k=0}^\infty c_kp^k\]
	where each $c_k$ lives in $\{0,1,\ldots,p-1\}$. The condition on $a_n$ requires that $c_k=0$ for $k<n$. We now define $(b_\nu)_\nu$ as from the power series
	\[\frac1{p^n}\sum_{k=0}^\infty c_kp^k=\sum_{k=0}^\infty c_{n+k}p^k.\]
	In particular, for each $\nu\ge1$, we see that
	\[p^nb_\nu\equiv\sum_{k=0}^{\nu-1}c_{n+k}p^{n+k}\equiv\sum_{k=0}^{\nu+n-1}c_kp^k\equiv a_\nu\pmod{p^\nu},\]
	so we conclude that $(a_\nu)_\nu\in p^n\ZZ_p$.
\end{proof}
\begin{corollary}
	Fix a prime $p$. Then the ring $\ZZ_p$ has a unique maximal ideal, which is $p\ZZ_p$.
\end{corollary}
\begin{proof}
	We claim that all proper ideals are contained in $p\ZZ_p$, which will imply the claim. In fact, we will show the contraposition of this statement: we claim that any ideal not contained in $p\ZZ_p$ is not a proper ideal. Well, suppose that $I\subseteq\ZZ_p$ is an ideal not contained in $p\ZZ_p$. Then there is some $(a_\nu)_\nu\in I$ which is not in $p\ZZ_p$. Well, \Cref{lem:principal-ideals-zp} tells us that $p\ZZ_p$ exactly consists of the $(b_\nu)_\nu\in\ZZ_p$ such that $p\mid b_\nu$ for each $\nu\ge1$, so \Cref{prop:units-in-zp} tells us that $\ZZ_p\setminus p\ZZ_p=\ZZ_p^\times$. As such $(a_\nu)_\nu\in I$ is a unit, which implies $I=\ZZ_p$.
\end{proof}
Having a unique maximal ideal which is principal is incredibly nice, but in fact the ideal structure of $\ZZ_p$ is as nice as it could possibly be.
% \begin{lemma}
% 	Fix a prime $p$, and define the function $v\colon\ZZ_p\setminus\{0\}\to\ZZ$ by
% 	\[v((a_\nu)_\nu)\coloneqq\min\left\{\nu\ge1:a_\nu\not\equiv0\pmod{p^\nu}\right\}.\]
% \end{lemma}
\begin{proposition}
	Fix a prime $p$. Then the ring $\ZZ_p$ is a principal ideal domain. In fact, for any nonzero ideal $I\subseteq\ZZ_p$, there exists a nonnegative integer $n$ such that
	\[I=p^n\ZZ_p.\]
\end{proposition}
\begin{proof}
	Note that $\ZZ_p$ is already an integral domain by \Cref{lem:zp-int-domain}. Now, let $I\subseteq\ZZ_p$ be a nonzero ideal. The main point of the proof is the construction of $n$. Because $I$ is nonzero, there exists some nonzero $(a_\nu)_\nu\in I$, and being nonzero implies that there is some greatest integer $n((a_\nu)_\nu)$ such that $a_n\not\equiv0\pmod{p^n}$. Define $n$ to be the least such integer over all nonzero $(a_\nu)_\nu\in I$.

	It remains to show that $I=p^n\ZZ_p$. One inclusion is not so bad: for each nonzero $(a_\nu)_\nu\in I$, the construction of $n$ requires that $a_n\equiv0\pmod{p^n}$, so $(a_\nu)_\nu\in p^n\ZZ_p$ by \Cref{lem:principal-ideals-zp}. For the other inclusion, we must show that $p^n\in I$. Well, by the construction of $n$, we are promised some $(a_\nu)_\nu\in I$ such that $a_n\equiv0\pmod{p^n}$ but $a_{n+1}\not\equiv0\pmod{p^{n+1}}$. By \Cref{lem:principal-ideals-zp} again, we are able to write
	\[(a_\nu)_\nu=p^n(b_\nu)_\nu\]
	for some $(b_\nu)_\nu\in\ZZ_p$. We would like to cancel out the $(b_\nu)_\nu$ from both sides.
	
	Well, if $b_{n+1}\equiv0\pmod p$, then $p^nb_{n+1}\equiv0\pmod{p^{n+1}}$, which contradicts the construction of $(a_\nu)_\nu$. Thus, we must instead have $b_{n+1}\in\left(\ZZ/p^{n+1}\ZZ\right)^\times$, so $(b_\nu)_\nu$ is a unit, and we may find $(c_\nu)_\nu\in\ZZ_p$ such that $(b_\nu)_\nu\cdot(c_\nu)_\nu=1$. Finishing up, we see
	\[p^n=(a_\nu)_\nu\cdot(c_\nu)_\nu\]
	lives in $I$, which completes the proof.
\end{proof}

\subsection{Hensel's Lemma}
Hensel's lemma is a tool which allows one to quickly check if a polynomial equation has roots in the $p$-adics. To orient ourselves, we begin with an example.
\begin{example} \label{ex:sqrt-2-z7}
	We show that $x^2-2=0$ has a solution in $\ZZ_7$.
\end{example}
\begin{proof}
	We must construct some $a\in\ZZ_7$ such that $a^2-2=0$, so we need to find a sequence $(a_\nu)_{\nu\in\NN}$ such that
	\[a_\nu^2\equiv2\pmod{7^\nu}\]
	for each $\nu\ge0$ and such that $a_{\nu+1}\equiv a_\nu\pmod{7^\nu}$. Well, we begin with $a_0\coloneqq0$ and $a_1\coloneqq3$ and then proceed inductively.
	
	For example, to solve for $a_2$, we see that we would like to set $a_2=a_1+\varepsilon$ for some $\varepsilon\equiv0\pmod7$. But we want $a_2^2\equiv2\pmod{7^2}$, which requires
	\[2\equiv a_2^2\equiv a_1^2+2a_1\varepsilon\pmod{7^2},\]
	so we select $\varepsilon$ by
	\[\varepsilon\equiv\frac{2-a_1^2}{2a_1}\pmod{7^2}.\]
	Note we may divide by $a_1\not\equiv0\pmod7$, and this is $0\pmod7$ by the construction of $a_1$.

	In general, suppose for $n\ge1$ we have been handed $a_n\pmod{7^n}$ with $a_n^2\equiv2\pmod{7^n}$ and $a_n\equiv a_1\pmod7$, and we want to construct $a_{n+1}\pmod{7^{n+1}}$. Well, we would like for $a_{n+1}=a_n+\varepsilon$ where $\varepsilon\equiv0\pmod{7^n}$, from which we see we require
	\[2\equiv a_{n+1}^2\equiv a_n^2+2a_n\varepsilon\pmod{7^{n+1}},\]
	where we are noting that $\varepsilon\equiv0\pmod{7^{n+1}}$. This rearranges to
	\[\varepsilon\equiv\frac{2-a_n^2}{2a_n}\pmod{7^{n+1}}.\]
	To see that this makes sense, we note that $a_n\equiv a_1\not\equiv0\pmod7$, so the denominator here is legal. Also, the numerator is $0\pmod{7^n}$ by hypothesis on $a_n$. So we see that we can construct $\varepsilon$ and hence $a_{n+1}\equiv a_n+\varepsilon$ will do.
\end{proof}
This example generalizes almost immediately to the following result.
\begin{proposition}[Hensel] \label{prop:hensel}
	Fix a prime $p$ and a polynomial $f\in\ZZ_p[x]$. Suppose we have some $a\in\ZZ_p$ such that $f(a)\equiv0\pmod p$ but $f'(a)\not\equiv0\pmod p$. Then there is a unique $\alpha\in\ZZ_p$ such that $\alpha\equiv a\pmod p$ and $f(\alpha)=0$.
\end{proposition}
Here, $f'$ denotes the derivative of a polynomial, which we can formally define via the power rule.
\begin{proof}
	We follow the proof of \Cref{ex:sqrt-2-z7}. We show by induction that there is a unique $a_n\pmod{p^n}$ such that $a_n\equiv a\pmod p$ and $f(a_n)\equiv0\pmod{p^n}$. For $n=0$, there is nothing to check. For $n=1$, we take $a_1\equiv a\pmod p$.

	It remains to do the inductive step. We are given $a_n\pmod{p^n}$ for $n\ge1$, and we must show that there is a unique $a_{n+1}\pmod{p^{n+1}}$ such that $a_{n+1}\equiv a\pmod p$ and $f(a_{n+1})\equiv0\pmod{p^{n+1}}$. Well, by the inductive step, we know that we must have $a_{n+1}\equiv a_n\pmod{p^n}$, so we will need to set $a_{n+1}=a_n+\varepsilon$ where $\varepsilon\equiv0\pmod{p^n}$. It remains to show that $a_{n+1}$ can be uniquely constructed from here.
	
	We would now like to approximate
	\[f(a_n+\varepsilon)\stackrel?\equiv0\pmod{p^{n+1}}.\]
	For example, one approach would be to use a Taylor expansion, writing
	\[f(a+\varepsilon)=\sum_{d=0}^\infty\frac{f^{(d)}(a)}{d!}\varepsilon^d,\]
	which makes sense in the fraction field $\QQ_p$ of $\ZZ_p$. One can show that the above equality works purely formally; notably, it is a finite sum because $f^{(d)}(a)=0$ for $d>\deg f$. However, this expansion is a little rough to bound with because of the presence of denominators---we have to move to $\QQ_p$ for it to make sense, after all. Instead, we pick up the following lemma which allows us to use the first-order approximation given by the Taylor expansion but avoiding the denominators.
	\begin{lemma} \label{lem:first-order-taylor}
		Fix a ring $R$ and a polynomial $f\in R[x]$. Then there is a polynomial $g\in R[x,\varepsilon]$ such that
		\[f(x+\varepsilon)=f(x)+f'(x)\varepsilon+\varepsilon^2g(x,\varepsilon).\]
		Here, $\varepsilon$ is being thought of as a formal variable.
	\end{lemma}
	\begin{proof}
		Write $f(x)=\sum_{d=0}^na_dx^d$ where $a_d\in R$ for each $d$. The main point is to show the result for the monomials $x^d$, where the result follows from using the binomial theorem to expand
		\begin{align*}
			(x+\varepsilon)^d &= x^d+dx^{d-1}\varepsilon+\sum_{k=2}^d\binom dkx^{d-k}\varepsilon^k \\
			&= x^d+dx^{d-1}\varepsilon+\varepsilon^2\underbrace{\sum_{k=2}^d\binom dkx^{d-k}\varepsilon^{k-2}}_{g_d(x,\varepsilon)\coloneqq}.
		\end{align*}
		Note this works even for $d=0$ as long as we simply ignore the higher-order terms. From here, we take $R$-linear combinations, writing
		\begin{align*}
			f(x+\varepsilon) &= \sum_{d=0}^na_d(x+\varepsilon)^d \\
			&= \sum_{d=0}^na_d\left(x^d+dx^{d-1}\varepsilon+\varepsilon^2g_d(x,\varepsilon)\right) \\
			&= \underbrace{\sum_{d=0}^na_dx^d}_{f(x)}+\Bigg(\underbrace{\sum_{d=0}a_ddx^{d-1}}_{f'(x)}\Bigg)\varepsilon+\varepsilon^2\underbrace{\sum_{d=0}^na_dg_d(x,\varepsilon)}_{g(x,\varepsilon)\coloneqq},
		\end{align*}
		which is what we wanted.
	\end{proof}
	Now, \Cref{lem:first-order-taylor} provides us with $g\in\ZZ_p[x,\varepsilon]$ such that
	\[f(a_n+\varepsilon)=f(a_n)+f'(a_n)\varepsilon+\varepsilon^2g(x,\varepsilon).\]
	Reducing$\pmod{p^{n+1}}$, we note that $\varepsilon^2\equiv0\pmod{p^{n+1}}$ because $n\ge1$, so we want
	\[0\stackrel?\equiv f(a_n)+f'(a_n)\varepsilon\pmod{p^{n+1}},\]
	which rearranges to
	\[\varepsilon\equiv-\frac{f(a_n)}{f'(a_n)}\pmod{p^{n+1}}.\]
	This fraction makes sense because $f'(a_n)\equiv f'(a)\not\equiv0\pmod p$---note that this is the only point in the proof where we have used the hypothesis that $f'(a)\not\equiv0\pmod p$! Additionally, we note that $f(a_n)\equiv0\pmod{p^n}$, so $\varepsilon\equiv0\pmod{p^n}$ as needed. So we have uniquely constructed a working $\varepsilon$, which provides a uniquely constructed $a_{n+1}$, as needed.
\end{proof}
Let's give a few applications.
\begin{corollary}
	Fix a prime $p$. For each nonzero residue $a\pmod p$, there is a unique $\zeta\in\ZZ_p$ such that $\zeta\equiv a\pmod p$ and $\zeta^{p-1}=1$.
\end{corollary}
\begin{proof}
	We apply \Cref{prop:hensel} to $f(x)\coloneqq x^{p-1}-1$. Indeed, for each nonzero residue $a\pmod p$, we see that $f(a)=a^{p-1}-1\equiv0\pmod p$ but $f'(a)=(p-1)a^{p-2}\not\equiv0\pmod p$.
\end{proof}
\begin{corollary} \label{cor:sqrt-zp}
	Fix an odd prime $p$. For each $\alpha\in\ZZ_p$ such that $\alpha\pmod p$ is a quadratic residue, there is $\beta\in\ZZ_p$ such that $\beta^2=\alpha$.
\end{corollary}
\begin{proof}
	We apply \Cref{prop:hensel} to $f(x)\coloneqq x^2-\alpha$. Indeed, $\alpha\pmod p$ is a quadratic residue, so we can find some nonzero $b\pmod p$ such that $b^2\equiv\alpha\pmod p$ so that $f(b)\equiv0\pmod p$; however, $f'(b)=2b\not\equiv0\pmod p$.
\end{proof}
\begin{example}
	Fix an odd prime $p$. For $a\in\ZZ$, if $a\pmod p$ is a quadratic residue, then we claim
	\[x^2\equiv a\pmod{p^n}\]
	has a solution for each nonnegative integer $n$. Indeed, \Cref{cor:sqrt-zp} promises $\beta\in\ZZ_p$ such that $\beta^2=a$, so $\beta\pmod{p^n}$ will do for each $n$.
\end{example}

\subsection{Newton's Method}
It will be worth our time to more closely examine the proof of \Cref{prop:hensel}. Approximately speaking, the proof is doing a $p$-adic version of ``Newton's method'' to find roots of a polynomial. Reading the proof over, it suggests taking some $a\in\ZZ_p$ such that $f(a)\approx0$ and then replacing $a$ with
\[a-\frac{f(a)}{f'(a)}\]
iteratively to eventually produce a root. Indeed, we approximated $f(a+\varepsilon)\approx f(a)+f'(a)\varepsilon$ and then found that we want to set $\varepsilon=-f(a)/f'(a)$ in order to make this vanish. Geometrically, what is happening is that we are pretending that $f$ locally looks its tangent line at $x=a$ and then finding the root of the tangent line
\[y-f(a)=f'(a)(x-a)\]
to produce a better approximate root. Here is the image for this.
\begin{center}
	\begin{asy}
		import graph;
		unitsize(4cm);
		real f(real x)
		{
			return x*x*x - x + 0.1;
		}
		draw((-0.6,0)--(2,0)); draw((0,-0.6)--(0,2));
		draw(graph(f,-0.6,1.5), blue);

		real a0 = 1.3;
		real a1 = a0 - (a0*a0*a0 - a0) / (3*a0*a0 - 1);

		pair v = (a0, f(a0)) - (a1, 0);
		draw((a1, 0) -- (a1, 0) + 1.8*v, red);

		draw((a0,0) -- (a0,f(a0)), dashed);
		draw((a1,0) -- (a1,f(a1)), dashed);

		dot("$a$", (a0,0), S); dot("$x_1$", (a1, 0), S);

		dot("$(a,f(a))$", (a0, f(a0)), WNW);
		dot("$(a',f(a'))$", (a1, f(a1)), WNW);

		label("$y-f(a)=f'(a)(x-a)$", (a1,0) + 1.5*v, W, red);
	\end{asy}
\end{center}
Now, a close reading of the proof of \Cref{prop:hensel} is able to show that this $p$-adic version of Newton's method works under only mild hypotheses.
\begin{lemma} \label{lem:newton-step}
	Fix a prime $p$ and a polynomial $f\in\ZZ_p[x]$. Suppose that we have some $a\in\ZZ_p$ and integers $N\ge0$ and $k>0$ such that the following hold.
	\begin{itemize}
		\item $f(a)\equiv0\pmod{p^{2N+k}}$,
		\item $f'(a)\equiv0\pmod{p^{N}}$, but
		\item $f'(a)\not\equiv0\pmod{p^{N+1}}$.
	\end{itemize}
	Then $a'\coloneqq a-f(a)/f'(a)$ has $f(a')\equiv0\pmod{p^{2N+2k}}$ and $a'\equiv a\pmod{p^{N+k}}$. In fact, $a'\pmod{p^{N+2k}}$ is uniquely determined by these constraints.
\end{lemma}
Intuitively, the conditions are requiring that $f(a)$ vanishes more than ``twice as much'' as $f'(a)$ in order to iterate. For example, at $N=0$, we are merely requiring that $f'(a)\not\equiv0\pmod p$, which is the condition appearing in \Cref{prop:hensel}.
\begin{proof}
	We imitate the inductive step of \Cref{prop:hensel}. For the uniqueness, we are required to choose $a'$ so that $a'=a+\varepsilon$ where $\varepsilon\equiv0\pmod{p^{N+k}}$, and we would like for $\varepsilon$ to be unique$\pmod{p^{N+2k}}$. Well, we use \Cref{lem:first-order-taylor} again to find $g$ with coefficients in $\ZZ_p$ such that
	\[f(a+\varepsilon)=f(a)+f'(a)\varepsilon+\varepsilon^2g(x,\varepsilon),\]
	meaning that we would like for
	\[f(a)+f'(a)\varepsilon\equiv0\pmod{p^{2N+2k}}\]
	because $\varepsilon^2\equiv0\pmod{p^{2N+2k}}$. This is now equivalent to
	\[\varepsilon\equiv-\frac{f(a)}{f'(a)}\pmod{p^{N+2k}},\]
	which we note is indeed $0\pmod{p^{N+k}}$ by using the conditions on $f(a)$ and $f'(a)$. So $\varepsilon=-f(a)/f'(a)$ will do, and it is unique$\pmod{p^{N+2k}}$, as required.
\end{proof}
\begin{proposition} \label{prop:hensel-by-newton}
	Fix a prime $p$ and a polynomial $f\in\ZZ_p[x]$. Suppose that we have some $a_0\in\ZZ_p$ and integers $N\ge0$ and $k>0$ such that the following hold.
	\begin{itemize}
		\item $f(a_0)\equiv0\pmod{p^{2N+k}}$.
		\item $f'(a_0)\equiv0\pmod{p^{N}}$.
		\item $f'(a_0)\not\equiv0\pmod{p^{N+1}}$.
	\end{itemize}
	Then the sequence $\{a_n\}_{n\in\NN}$ defined by $a_{n+1}\coloneqq a_n-f(a_n)/f'(a_n)$ has
	\[f(a_n)\equiv0\pmod{p^{2N+2^nk}}\qquad\text{and}\qquad a_{n+1}\equiv a_n\pmod{p^{N+2^nk}}\]
	for each $n\ge0$. This sequence then constructs a unique $\alpha\in\ZZ_p$ such that $f(\alpha)=0$ and $\alpha\equiv a_0\pmod{p^{N+1}}$.
\end{proposition}
\begin{proof}
	We inductively apply \Cref{lem:newton-step}. At $n=0$, there is nothing to do. For the induction, we note that we are given $a_n$ for $n\ge0$ satisfying the following.
	\begin{itemize}
		\item $f(a_0)\equiv0\pmod{p^{2N+2^nk}}$.
		\item Note that $a_n\equiv a_0\pmod{p^{N+1}}$ implies $f'(a_n)\equiv0\pmod{p^N}$ while $f'(a_n)\not\equiv0\pmod{p^{N+1}}$.
	\end{itemize}
	From here, \Cref{lem:newton-step} promises $a_{n+1}=a_n-f(a_n)/f'(a_n)$ will have $f(a_{n+1})\equiv0\pmod{p^{2N+2^{n+1}k}}$ and $a_{n+1}\equiv a_n\pmod{p^{N+2^nk}}$, and we know that $a_{n+1}\pmod{p^{N+2^{n+1}k}}$ is uniquely determined by these constraints. This has shown the first claim.

	It remains to prove the last sentence. The fact that $a_{n+1}\equiv a_n\pmod{p^{N+2^nk}}$ for each $n\ge0$ promises that there is some $\alpha\in\ZZ_p$ such that $\alpha\equiv a_n\pmod{p^{N+2^nk}}$ for each $n\ge0$. To see that this $\alpha$ is unique, we claim that $f(\alpha)=0$ and $\alpha\equiv a_n\pmod{p^{N+1}}$ requires that $\alpha\equiv a_n\pmod{p^{N+2^n}}$ for each $n\ge0$, which will complete the proof upon sending $n\to\infty$. For this, we use induction; the case of $n=0$ has nothing to check. Then for the induction, we note that \Cref{lem:newton-step} shows that $a_{n+1}\pmod{p^{N+2^{n+1}}}$ is uniquely determined by having
	\[f(a_{n+1})\equiv0\pmod{p^{2N+2^{n+1}}}\qquad\text{and}\qquad a_{n+1}\equiv a_n\pmod{p^{N+2^n}}.\]
	Well, $f(\alpha)=0$ and $\alpha\equiv a_n\pmod{p^{N+2^n}}$ already, so the uniqueness requires $\alpha\equiv a_{n+1}\pmod{p^{N+2^{n+1}}}$ too.
\end{proof}
\begin{remark}
	The reason we have stated \Cref{prop:hensel-by-newton} so carefully is to point out that even though the proof of \Cref{prop:hensel} constructs the unique root of $f$ digit-by-digit, the exact same method in fact provides us with a way to converge to the root doubling the amount of precision we have at each step! This is an amazing speed-up!
\end{remark}
We now close with a final example, proving \Cref{prop:squares-z2}.
\ztwosquares*
\begin{proof}
	There is nothing to show for (b) implies (a). To show that (a) implies (b), we use \Cref{prop:hensel-by-newton} with the polynomial $f(x)=x^2-a$. Indeed, we see that $f(1)\equiv0\pmod8$ and $f'(1)\not\equiv0\pmod4$, so we are given some $\alpha\in\ZZ_2$ such that $\alpha^2=a$. Thus, $\alpha\pmod{2^\nu}$ for each $\nu\ge0$ will have $\alpha^2\equiv a\pmod{2^\nu}$.
\end{proof}

\subsection{Problems}
Do ten points worth of the following exercises and then complete the survey.
\begin{prob}[2 points]
	Fix a prime $p$. Show that
	\[\sum_{n=0}^\infty(n-1)n!=-1\]
	in $\ZZ_p$ by showing that both sides agree$\pmod{p^n\nu}$ for any $\nu\ge1$.
\end{prob}
\begin{prob}[2 points]
	Compute a sequence of integers $c_1,c_2,c_3,\ldots\in\{0,1,\ldots,10\}$ such that
	\[-\frac23=\sum_{k=0}^\infty c_k11^k\]
	in $\ZZ_{11}$.
\end{prob}
\begin{prob}[3 points]
	Find an integer $a$ such that $a^2\equiv-1\pmod{125}$.
\end{prob}
\begin{prob}[3 points]
	Fix a prime $p$, and choose a subset $S\subseteq\ZZ$ containing exactly one representative of each class in $\ZZ/p\ZZ$. For any $a\in\ZZ_p$, show that there is a unique sequence $(c_\nu)_{\nu=0}^\infty\in S^\NN$ such that
	\[a=\Bigg(\sum_{k=0}^{\nu-1}c_kp^k\Bigg)_{\nu=0}^\infty.\]
\end{prob}
\begin{prob}[3 points]
	Fix some $a\in\ZZ_2$. Show that the following are equivalent.
	\begin{listalph}
		\item There is some $b\in\ZZ_2$ such that $a^2=b$.
		\item There is a positive integer $k$ and integer $a'\equiv1\pmod8$ such that
		\[a\equiv 4^ka'\pmod{2^{2k+3}}.\]
	\end{listalph}
\end{prob}
\begin{prob}[4 points]
	Fix distinct primes $p>q$.
	\begin{listalph}
		\item Show that $x^p-x$ has at most $q$ roots in $\ZZ/q\ZZ$ and hence at most $q$ roots in $\ZZ_q$.
		\item Show that $x^p-x$ has at most $p$ roots in $\ZZ/p\ZZ$ and hence exactly $p$ roots in $\ZZ_p$.
		\item Using (a) and (b), conclude that the rings $\ZZ_p$ and $\ZZ_q$ are not isomorphic.
	\end{listalph}
	Similar logic can show that the fraction fields $\QQ_p$ and $\QQ_q$ of $\ZZ_p$ and $\ZZ_q$ are not isomorphic for distinct primes $p>q$.
\end{prob}
\begin{prob}[4 points] \label{prob:223-in-z3}
	Show that $x^2-223y^2=-3$ has a solution in $\ZZ_3$ by showing that there exists $y\in\ZZ_3$ such that $y^2=4/223$. Use \Cref{prob:norm-223} to conclude that the equation
	\[x^2-223y^2\equiv-3\pmod n\]
	has a solution for any positive integer $n$.
\end{prob}
\begin{prob}[6 points]
	Write a Python program which takes as input some (small) odd prime $p$, an integer $a\in\ZZ$ such that $a\pmod p$ is a quadratic residue, and a precision level $n$, and then outputs some integer $b_n$ such that $b_n^2\equiv a\pmod{p^n}$. Your program should use Hensel's lemma; as an example, compute a square root of $-1\pmod{5^{100}}$.
\end{prob}
\begin{prob}[0 points]
	Please rate the speed of the following lectures, from ``much too slow'' to ``much too fast.''
	\begin{itemize}
		\item November 13: The Group Law
		\item November 15: Weierstrass Form
		\item November 17: Explicit Group Laws
	\end{itemize}
	Please also rate the difficulty of the problems on the homework you solved.
\end{prob}

% show Z_p is a PID with a unique maximal ideal

% define Z_p algebraically, and get Q_p as a quotient
% show that Z is dense in Z_p, but Z_p has more elements than Z
% exhibit the absolute value
% give redefinition of Qp
% ostrowski
% product formula

% 10-adics?
% power series? mahler expansions?
% other fields? I don't know if there are applications ...

% strassman's theorem to solve x^3 - 2y^3 = 1

\end{document}