% !TEX root = ../notes.tex

\documentclass[../notes.tex]{subfiles}

\begin{document}

\section{February 8}

Last class we showed that every curve can be embedded into $\PP^3_k$.
\begin{remark}
	We will not show this, but one can show with a little more work that any curve is birational to a curve in the plane with at worst nodes as singularities.
\end{remark}

\subsection{The Canonical Embedding}
Throughout, $X$ is a $k$-curve, where $k$ is algebraically closed. We will set $g\coloneqq g(X)$ and let $K$ denote the canonical divisor.
\begin{example}
	Notably, $g(X)=0$ forces $\dim|K|=-1$, so $|K|$ is empty. (After all, $X\cong\PP^1_k$.)
\end{example}
\begin{example}
	If $g(X)=1$, then we are looking at $\dim|K|=0$, and here $\deg K=0$, so we are just looking at the mapping from $X$ to a point.
\end{example}
In higher genus, things get more interesting.
\begin{lemma} \label{lem:k-is-base-point-free}
	Fix a $k$-curve $X$ with $g(X)\ge2$. Then the canonical divisor $K$ is base-point-free.
\end{lemma}
\begin{proof}
	By \Cref{prop:div-to-proj-embed}, it suffices to show that $\ell(K-P)=\ell(K)-1=g-1$ for any $P\in X$. Now, because $g(X)\ge2>0$, we know that $X\not\cong\PP^1_k$, so $\dim|P|=0$ is forced. Now, using \Cref{thm:rh}, we solve
	\[\ell(P)=\ell(K-P)+\deg(P)+(1-g),\]
	so $\ell(K-P)=g-1$, which is what we wanted.\todo{What?}
\end{proof}
In our study of curves, the following curves will make a somewhat large class.
\begin{definition}[hyperelliptic]
	A $k$-curve $X$ is \textit{hyperelliptic} if and only if it admits a degree-$2$ map to $\PP^1_k$.
\end{definition}
\begin{remark}
	Given any divisor $D$ on $X$ with $\ell(D)=2$ and $\deg D=2$, one gets a rational map $X\dashrightarrow\PP^1_k$ determined by $D$. (One can show that all such divisors are base-point-free using \Cref{prop:div-to-proj-embed}.) Notably, $D$ is linearly equivalent to an effective divisor.
\end{remark}
\begin{remark}
	If $g(X)=2$, then $X$ is hyperelliptic. This was shown on the homework. In brief, $K$ is base-point-free and has degree $2$, so it determines a degree-$2$ map $X\to\PP^1_k$.
\end{remark}
One can improve \Cref{lem:k-is-base-point-free} as follows.
\begin{proposition} \label{prop:k-is-very-ample}
	Fix a $k$-curve $X$ with $g(X)\ge2$. Then $K$ is very ample if and only if $X$ is not hyperelliptic.
\end{proposition}
\begin{proof}
	By \Cref{prop:div-to-proj-embed}, we are interested in the condition
	\[\ell(K-P-Q)=\ell(K)-2\stackrel?=g-2\]
	for any $P,Q\in X$. As such, by \Cref{thm:rh}, we compute
	\[\ell(P+Q)=\ell(K-P-Q)+\deg(P+Q)+(1-g)=\ell(K-P-Q)+3-g,\]
	so $\ell(K-P-Q)=g-2$ is equivalent to
	\[\ell(P+Q)=1.\]
	Now, if $X$ is hyperelliptic, then one can find a divisor $D$ with $\dim|D|=1$ and $\deg D=2$, so $D$ linearly equivalent to an effective divisor which looks like $P+Q$. But then $\ell(P+Q)=2>1$, which is a problem. In the other direction, if $X$ is not hyperelliptic, then each effective divisor $D$ with degree $2$ must have $D=P+Q$, which forces $\ell(P+Q)<2$ to retain being not hyperelliptic.
\end{proof}
Thus, we are interested in the embedding induced by this canonical divisor.
\begin{definition}[canonical morphism]
	Fix a $k$-curve $X$ of genus $g(X)\ge2$. Then the canonical divisor $K$ determines the \textit{canonical morphism} $X\to\PP^{g-1}_k$ by \Cref{lem:k-is-base-point-free}. If $X$ is not hyperelliptic curves, then we in fact get a \textit{canonical embedding} by \Cref{prop:k-is-very-ample}.
\end{definition}
\begin{remark}
	If $\deg\mc L=d$, and $\mc L$ is very ample, then $\mc L$ grants a closed embedding of $X$ to $\PP^{r-1}_k$ where $r\coloneqq\dim_k\Gamma(X,\mc L)$ (using the canonical embedding), and this embedding retains $X$ as a curve of degree $d$.
\end{remark}

\subsection{Hyperelliptic Curves}
We will spend the rest of class understanding hyperelliptic curves. The following is our statement.
\begin{theorem}
	Fix a hyperelliptic curve $X$ of genus $g(X)\ge2$.
	\begin{listalph}
		\item Then $X$ has a unique divisor $g^1_2$ yielding the double-cover $\pi\colon X\to\PP^1_k$.
		\item The canonical morphism $f\colon X\to\PP^{g-1}$ can be written as
		\[X\to\PP^1_k\to\PP^{g-1},\]
		where the last map is the $(g-1)$-uple embedding.
		\item Every effective canonical divisor can be written as the sum of $(g-1)$ effective divisors linearly equivalent to $g^1_2$.
	\end{listalph}
\end{theorem}
\begin{proof}
	For brevity, let $X'$ be the image $f(X)$, and we fix some $g^1_2$ on $X$ yielding a double-cover $X\to\PP^1_k$.
	
	Now, for any effective divisor $P+Q$ linearly equivalent to $g^1_2$, then we claim $Q$ is a base-point of $K-P$: indeed, we know $\ell(K-P)=g-1$ because $K$ is base-point-free by \Cref{lem:k-is-base-point-free}, but $K$ is not very ample witnessed by $P$ and $Q$ by the proof of \Cref{prop:k-is-very-ample}, so $\ell(K-P-Q)=\ell(K-P)$ is forced. Thus, $Q$ is in fact a base-point.

	We now have two cases.
	\begin{itemize}
		\item In the case where $P\ne Q$, we see that $K$ does not separate the points $P$ and $Q$. Explicitly, we see that $s\in\Gamma(X,\OO_X(K-P))$ is equivalent to $s\in\Gamma(X,\OO_X(K-P-Q))$ is equivalent to $s\in\Gamma(X,\OO_X(K-Q))$ (by symmetry). In other words, we are saying that any divisor $K+\op{div}(s)$ which retains $P$ in its support will also retain $Q$ in its support.
		
		It follows that $f(P)=f(Q)$: otherwise $f(P)\ne f(Q)$ lets us separate the two points in $\PP^{g-1}_k$ and hence find a basis of $\Gamma(\PP^{g-1}_k,\OO_{\PP^{g-1}}(1))$ separating them. But then we can pull back this basis to $\Gamma(X,\OO_X)$ to find sections separating $P$ and $Q$.

		\item Otherwise, we in the case where $P=Q$, we are now given that $P$ is a base-point of $K-P$. Explicitly, we know that $s\in\Gamma(X,\OO_X(K-P))$ is equivalent to $s\in\Gamma(X,\OO_X(K-2P))$, which now means that $f$ does not separate tangent vectors at $P$---namely, the image of $\Gamma(X,\OO_X(K))$ does not generate $\mf m_P\mc O_X(K)_P/\mf m_P/\OO_X(K)_P$. However, $\OO_{\PP^{g-1}_k}(1)$ does separate these tangent spaces by hyperplanes, so $f$ is again not a closed embedding at the local ring at $P$.
	\end{itemize}
	By adjusting $P+Q$ appropriately in its linear equivalence class, we see that $f$ fails to be a closed embedding at infinitely many points, so $f$ is not a birational morphism.

	As such, $\mu\coloneqq\deg f$ is at least $2$. However, letting $d$ denote the degree of $X'\subseteq\PP^{g-1}_k$, we claim
	\[d\mu\stackrel?=2g-2.\]
	For this, we must understand $d=\deg X'$, which we note is the number of intersections (counted with multiplicity) of a hyperplane section of $\PP^{g-1}_k$ intersecting $Y=X'\setminus\op{Sing}X'$, where $\op{Sing}X'$ is the singular locus. Letting $H\cap Y$ denote such an intersection (with $d$ points and hence degree $d$ as a divisor), we restrict $f\colon X\to X'$ to $f\colon U\to\widetilde{X'}$ where $U\coloneqq f^{-1}(Y)$, granting
	\[\deg (f|_U)^*(H\cap Y)=(\deg f)d=\mu d\]
	by explicitly writing down what does $f^*$ does on points on the finite morphism of smooth curves $f|_{\widetilde{X'}}$. We will finish this proof next class.
\end{proof}

\end{document}