% !TEX root = ../notes.tex

\documentclass[../notes.tex]{subfiles}

\begin{document}

We spent most of the class completing the proof of \Cref{thm:serre}. I have edited into those notes for continuity reasons.
\begin{remark}
	Note the identity map $A\to\Gamma(X,\OO_X)$ induces a map $\varphi\colon X\to\Spec A$ via the adjunction. So after we produce the sections $(f_1,\ldots,f_r)$ with $X_{f_i}$ affine, it might appear that we are done because we might be able to glue the $X_{f_i}\cong\Spec A_{f_i}$ into making $\varphi$ an isomorphism. But this does not work: indeed, $\varphi$ need not even be surjective!
\end{remark}

\subsection{More Cohomology on Affine Schemes}
Let's see an application of some of the work we've done.
\begin{corollary} \label{cor:embed-to-qcoh-flasque}
	Fix a Noetherian scheme $X$. Any quasicoherent sheaf $\mc F$ on $X$ can be embedded into a flasque quasicoherent sheaf.
\end{corollary}
\begin{proof}
	The point is to reduce to the affine case and then use \Cref{prop:inj-to-flasque-qcoh}. Let $\{U_i\}_{i=1}^n$ be a finite affine open cover of $X$, where $U_i=\Spec A_i$. Because $\mc F$ is quasicoherent, we can find an $A_i$-module $M_i$ such that $\mc F|_{U_i}\cong M_i$, and then we may embed $M_i$ into an injective $A_i$-module $I_i$. So we have injections $\mc F|_{U_i}\to\widetilde I_i$ on $U_i$, so we can glue these into a map
	\[\mc F\to\bigoplus_{i=1}^n(j_i)_*\widetilde I_i,\]
	where $j_i\colon U_i\to X$ is the inclusion. Notably, $(j_i)_*$ sends quasicoherent sheaves to quasicoherent sheaves (note $X$ is Noetherian), so $\bigoplus_{i=1}^n(j_i)_*\widetilde I_i$ is still quasicoherent, and the above map is injective because we can check injectivity on stalks and so on the affine open cover $\{U_i\}_{i=1}^n$. Lastly, $(j_i)_*$ sends flasque sheaves to flasque sheaves, so our sum is still flasque.
\end{proof}

\end{document}