% !TEX root = ../notes.tex

\documentclass[../notes.tex]{subfiles}

\begin{document}

\section{February 18}
Here we go.

\subsection{How to Grade}
We are going to prove \Cref{thm:better-hodge} by finding new structure on $X^{\mathrm{dR}}$.
\begin{notation}
	We define $\mathbb G_m^\sharp$ to be the group scheme with ring given by
	\[\Spec k\left[t,1/t,\{(t-1)^n/n!\}_n\right].\]
\end{notation}
\begin{remark}
	There is a canonical map $k[t,1/t]\to k[t,1/t,\{(t-1)^n/n!\}_n]$, but note that it factors through
	\[\frac{k[t,1/t]}{(t-1)^p}\]
	because $(t-1)^p$ is divisible by $p$ in the target. Thus, the canonical map $\mathbb G_m^\sharp\to\mathbb G_m$ factors through $\mu_p$.
\end{remark}
\begin{remark}
	There is a map $\log\colon\mathbb G_m^\sharp\to\mathbb G_a$ given by
	\[\log t\coloneqq\sum_{n\ge1}(-1)^{n+1}\frac{(t-1)^n}{n},\]
	and one can calculate that $\ker\log=\mu_p$. (Note that this sum is finite because $(t-1)^n/n=(t-1)^n/n!\cdot(n-1)!$, and $(n-1)!$ is eventually divisible by $p$.)
\end{remark}
\begin{theorem} \label{thm:dr-structure}
	Fix a smooth variety $X$ over a perfect field $k$ of positive characteristic $p$. Given a smooth lift $\widetilde X$ over $W_2(k)$, there is a natural action by $\mathbb G_m^\sharp$ on $X^{\mathrm{dR}}$.
\end{theorem}
\begin{proof}[Proof that \Cref{thm:dr-structure} implies \Cref{thm:better-hodge}]
	We will only explain how to produce the grading; the requirement in low degrees follows from the construction of the action. Now, an action of $\mathbb G_m^\sharp$ on $X^{\mathrm{dR}}$ amounts to an action of $\mathbb G_m^\sharp$ on
	\[\mathrm H^\bullet_{\mathrm{dR}}(X/k)=\op{Ext}^\bullet_{\mathrm{QCoh}(X^{\mathrm{dR}})}(\OO_{X^{\mathrm{dR}}},\OO_{X^{\mathrm{dR}}})\]
	because $\op{QCoh}(X^{\mathrm{dR}})$ is the category of quasi-nilpotent $\mc D$-modules by \Cref{rem:de-rham-by-quasinilpotent}.

	Now, a $\mathbb G_m$-action on a vector space is the data of a $\ZZ$-grading (by taking the eigenvalues of the $\mathbb G_m$-action). Thus, we may hope that our $\mathbb G_m^\sharp$-action can produce a similar grading. Thus, there is a canonical copy of $\mu_p$ in $\mathbb G_m^\sharp$, which then acts on $\mathrm H^\bullet_{\mathrm{dR}}(X/k)$. But now an action of $\mu_p$ on $V$ corresponds to a $\ZZ/p\ZZ$-grading given by taking the eigenspaces $V_i\coloneqq\{v\in V:tv=t^v\text{ for }t\in\mu_p\}$. We have thus exhibited the grading!
\end{proof}
\begin{remark}
	It turns out that $\mathbb G_m^\sharp\cong\mu_p\times\mathbb G_a^\sharp$. Thus, there is an interesting action by $\mathbb G_a^\sharp$ on the degree pieces.
\end{remark}
For motivation, note that \Cref{thm:dr-structure} can be refined if we give ourselves more lifting.
\begin{proposition} \label{prop:dr-structure-frob-lift}
	Fix a smooth variety $X$ over a perfect field $k$ of positive characteristic $p$. Given a smooth lift $\widetilde X$ over $W_2(k)$ and a lift $\widetilde F\colon\widetilde X\to\widetilde X^{(1)}$ of the Frobenius, there is a natural action of $\mathbb G_m$ on $X^{\mathrm{dR}}$.
\end{proposition}
\begin{remark}
	As before, a natural action of $\mathbb G_m$ on $X^{\mathrm{dR}}$ induces a $\ZZ$-grading on the de Rham cohomology.
\end{remark}
One may hope to use \Cref{prop:dr-structure-frob-lift} by working with $\AA^1$ and then passing to $X$ using the $k$-algebra stack structure. This is not possible because there is no Frobenius lift of $\AA^1$ on $W_2(k)$ which commutes with the structure as an additive group scheme: indeed, even for $k=\FF_p$, there is no polynomial $f(x)\in\ZZ/p^2\ZZ[x]$ such that $f(x+y)=f(x)+f(y)$ and $f(x)\equiv x^p\pmod p$. Thus, we cannot hope to use some ``canonical'' Frobenius lifts and also reduce everything to the ring stack structure of $\AA^1$: one needs to use the data of the lift to $W_2(k)$ to produce the Frobenius lift.

Anyway, let's prove \Cref{thm:dr-structure} in the case of $\AA^1$.
\begin{lemma}
	The kernel of the projection $\mathbb G_a^\sharp\to\mathbb G_a$ is given by $\mathbb G_a^{\sharp,(1)}$.
\end{lemma}
\begin{proof}
	The idea is that the map $\mathbb G_a^\sharp\to\mathbb G_a$ factors through $\alpha_p$ because $t^p\in k[t]$ is divisible by $p$ in the divided power ring, which is the kernel of $F\colon\mathbb G_a\to\mathbb G_a^{(1)}$.
\end{proof}
\begin{example}
	We exhibit a $\mathbb G_m^\sharp$-action on $\AA^{1,\mathrm{dR}}=\AA^1/\mathbb G_a^\sharp$. Indeed, $\mathbb G_m^\sharp$ acts on $\mathbb G_a^{\sharp,(1)}$ simply by scaling $\lambda\colon t^p/p\mapsto\lambda t^p/p$. This then embeds into $\mathbb G_a^\sharp$ as the kernel down to $\mathbb G_a$. The action follows.
\end{example}

\subsection{Witt Vectors}
We now move to the general case. Motivated by \Cref{prop:dr-structure-frob-lift}, we try to find a ``universal'' algebra with a Frobenius lift, which is the Witt vectors.
\begin{lemma}
	There is a unique flat $\ZZ_p$-algebra $\OO(W)$ with a Frobenius lift $F\colon \OO(W)\to \OO(W)$ such that $F(y)\equiv y^p\pmod p$ for all $y$ satisfying the following universal property: for every map $\ZZ_p[x]\to R$ with $R$ flat and admitting a Frobenius lift factors uniquely through $\OO(W)$.
\end{lemma}
\begin{proof}
	As one does in commutative algebra, we simply define
	\[\OO(W)\coloneqq\ZZ_p\left[x,F(x),\frac{F(x)-x^p}p,F\left(\frac{F(x)-x^p}p\right),\ldots\right],\]
	where $F(x)$ is viewed as a formal variable, and then one explicitly finds some generators and relations. There is some work to show that $W$ is flat, and it satisfies the universal property.
\end{proof}
\begin{definition}[Witt vectors]
	We define the \textit{Witt vector scheme} $W$ to be the scheme $\Spec\OO(W)$.
\end{definition}
\begin{proposition}
	There is a unique ring scheme structure on $W$ such that the map $W\to\AA^1$ is a ring map. In fact, $\OO(W)$ admits a presentation
	\[\OO(W)=\ZZ_p[a_0,a_1,\ldots]\]
	so that $F^n(a_0)=a_0^{p^n}+pa_1^{p^{n-1}}+\cdots+p^na_n$.
\end{proposition}
\begin{proof}
	Omitted.
\end{proof}
\begin{example}
	One has that $W(\FF_p)=\ZZ_p$.
\end{example}
\begin{remark}
	Here is how to work formally with $W(R)$: as a set, we know that $W(R)=\prod_{m\ge0}R$, and we must have that the ``ghost component'' map
	\[\prod_{m\ge0}R\to R\]
	given by $\mapsto a_0^{p^n}+\cdots+p^na_n$ is a ring map. For example, $(a+b)_0=a_0+b_0$, and $(a+b)_1=a_1+b_1+\frac{a_0^p+b_0^p-(a_0+b_0)^p}{p}$.
\end{remark}
\begin{remark}
	The Witt vectors $W$ admit a Frobenius lift, which continues to be a ring map.
\end{remark}
We are now ready to give some kind of Frobenius lift to $\AA^{1,\mathrm{dR}}$.
\begin{lemma} \label{lem:de-rham-frob-lift}
	The de Rham stack $\AA^{1,\mathrm{dR}}$ is isomorphic to $[W/pW]$, where $pW$ means that $W$ is acting additively on $W$ by multiplication by $p$.
\end{lemma}
\begin{proof}
	It turns out that the kernel of $F\colon W\to W$ is $\mathbb G_a^\sharp$. For example, $(Fa)_0=a_0^p+pa_1$, so $a_0$ has $p$th power divisible by $p$, which explains why divided powers may appear (and allows us to define a map $\mathbb G_a^\sharp\to\ker F$).
\end{proof}
\begin{proof}[Proof of \Cref{thm:dr-structure}]
	We would like a ring stack action of $\mathbb G_m^\sharp$ on $\AA^{1,\mathrm{dR}}$; this then formally goes to all $X^{\mathrm{dR}}$s. By \Cref{lem:de-rham-frob-lift}, we know that $\AA^{1,\mathrm{dR}}=[W/pW]$. Taking units, we know that $\mathbb G_m^\sharp$ is isomorphic to the kernel of $F\colon W^\times\to W^\times$. Thus, $\mathbb G_m^\sharp$ is able to rescale $W$, and it intertwines the multiplication-by-$p$ map, where the target $W$ has trivial action by $\mathbb G_m^\sharp$.

	Let's quickly explain why one may need a smooth lift of $X$. Suppose in fact that $X$ admits a lift $\widetilde X$ fully to $W(k)$. Then, working locally on $X$, we may assume that $\widetilde X$ is a locally complete intersection (e.g., the fiber of some $f\colon\AA^n\to\AA^m$ even over $W(k)$), and then the action on $X^{\mathrm{dR}}$ is found by pullback on ring stacks $f^{\mathrm{dR}}\colon\AA^{n,\mathrm{dR}}\to\AA^{m,\mathrm{dR}}$.
\end{proof}
\begin{remark}
	If we merely have a smooth lift to $W_2(k)$, then one has to redo the argument replacing $W$s with $W_2$s everywhere.
\end{remark}

\end{document}