% !TEX root = ../notes.tex

\documentclass[../notes.tex]{subfiles}

\begin{document}

\section{November 16}

Here we go.

\subsection{Integration Facts}
We continue our fact-collection.
\begin{lemma} \label{lem:almost-simple-int-are-dense}
	Fix a measure space $(X,\mc S,\mu)$ and a Banach space $B$. Further, fix a $\mu$-integrable function $f\colon X\to B$ with corresponding sequence mean Cauchy sequence $\{f_n\}_{n\in\NN}$ of simple $\mu$-integrable functions such that $f_n\to f$ in measure. Then $\norm{f-f_n}_1\to0$ as $n\to\infty$.
\end{lemma}
\begin{proof}
	This essentially follows directly from the definition of integration. Indeed, fix some $\varepsilon>0$. Our sequence is mean Cauchy, so choose some $N$ for which $m,n\ge N$ implies $\norm{f_m-f_n}_1<\varepsilon/2$.
	
	Now, for some fixed $m$, define $g_n\coloneqq f_m-f_n$ for each $n\in\NN$, which is a simple $\mu$-integrable function by \autoref{lem:simple-int-is-k-vec}, and we see $\{g_n\}_{n\in\NN}$ is mean Cauchy by \autoref{lem:linear-combo-mean-cauchy} with $g_n\to f_m-f$ in measure by \autoref{lem:linear-combo-in-measure}. Now, it follows from the proof that \autoref{lem:norm-int-function} that $\{\norm{g_n}\}_{n\in\NN}$ is still a mean Cauchy sequence of simple $\mu$-integrable functions such that $\norm{g_n}\to\norm{f_m-f}$ in measure, so
	\[\norm{f_m-f}_1=\int_X\norm{f_m-f}\,d\mu=\lim_{n\to\infty}\int_X\norm{g_n}\,d\mu=\lim_{n\to\infty}\norm{f_m-f_n}_1.\]
	(All the relevant functions are $\mu$-integrable by \autoref{lem:int-is-vec-space} and \autoref{lem:norm-int-function}.) Thus, taking $m\ge N$, we see $\norm{f_m-f_n}_1<\varepsilon/2$ for $n\ge N$, so
	\[\norm{f_m-f}_1=\lim_{n\to\infty}\norm{f_m-f_n}_1\le\varepsilon/2<\varepsilon.\]
	This completes the proof.
\end{proof}
The point of the above lemma is the following density result.
\begin{corollary} \label{cor:simple-int-are-dense}
	Fix a measure space $(X,\mc S,\mu)$ and a Banach space $B$. For any $\mu$-integrable function $f\colon X\to B$ and error $\varepsilon>0$, there is a simple $\mu$-integrable function $g\colon X\to B$ such that $\norm{f-g}_1<\varepsilon$.
\end{corollary}
\begin{proof}
	Because $f\colon X\to B$ is integrable, there is a mean Cauchy sequence $\{f_n\}_{n\in\NN}$ of simple $\mu$-integrable functions such that $f_n\to f$ in measure. But then \autoref{lem:almost-simple-int-are-dense} tells us that
	\[\lim_{n\to\infty}\norm{f-f_n}_1=0,\]
	so there is some $N$ such that $n\ge N$ implies $\norm{f-f_n}_1<\varepsilon$. Choosing any $n\ge N$ and setting $g\coloneqq f_n$ thus finishes.
\end{proof}
\begin{lemma} \label{lem:almost-support-l1}
	Fix a measure space $(X,\mc S,\mu)$ and a normed vector space $(B,\norm\cdot)$. Given a $\mu$-integrable function $f\colon X\to B$ and bound $\varepsilon>0$, there is some $F\subseteq X$ with $F\in\mc S$ and $\mu(F)<\infty$ such that
	\[\int_X\norm f1_{X\setminus F}\,d\mu<\varepsilon.\]
\end{lemma}
\begin{proof}
	Because $f\colon X\to B$ is $\mu$-integrable, we may choose some simple $\mu$-integrable function $g\colon X\to B$ such that $\norm{f-g}_1<\varepsilon$, where we are using \autoref{cor:simple-int-are-dense}. Now, choose $F\coloneqq g^{-1}(B\setminus\{0\})$, which is in $\mc S$ again using \autoref{lem:simple-measurable-pre-image}, and we note \autoref{lem:fin-additive-is-monotone} implies
	\[\mu(F)\le\mu\left(g^{-1}(B\setminus\{0\})\right),\]
	where $\mu\left(g^{-1}(B\setminus\{0\})\right)$ is finite by \autoref{rem:better-simp-int}. Thus, $\mu(F)<\infty$.
	
	It remains to compute $\int_X\norm f1_{X\setminus F}\,d\mu$. Well, we see $g(x)=0$ for $x\notin F$, so
	\[(\norm f1_{X\setminus F})(x)=(\norm{f-g}1_{X\setminus F})(x)\le\norm{f-g}(x)\]
	for each $x\in X$, so \autoref{lem:bound-ints} tells us
	\[\int_X\norm f1_{X\setminus F}\,d\mu\le\int_X\norm{f-g}\,d\mu=\norm{f-g}_1<\varepsilon,\]
	which is what we wanted.
\end{proof}
\begin{remark}
	The above result basically says that $f$ is almost supported on a set of finite measure.
\end{remark}
\begin{lemma} \label{lem:bound-meas-set}
	Fix a measure space $(X,\mc S,\mu)$. Given a $\mu$-integrable function $f\colon X\to\RR$, given $E\in\mc S$ with $f(x)\ge1_E(x)$ almost everywhere, then
	\[\mu(E)\le\int_Xf\,d\mu.\]
\end{lemma}
\begin{proof}
	The main difficulty here is that we don't actually know if $1_E$ is an integrable function at the outset.
	
	For convenience, we set $F\coloneqq f^{-1}(B\setminus\{0\})$. We claim that $F$ is contained in the countable union of sets of finite measure; this is annoying, so we will brief. Well, because $f\colon X\to\RR$ is $\mu$-integrable, we can find a mean Cauchy sequence $\{f_n\}_{n\in\NN}$ of simple $\mu$-integrable functions such that $f_n\to f$ in measure. Now, $g_n\to f$ almost everywhere (because $g_n\to f$ in measure), so there is some $N\in\mc S$ such that $\mu(N)=0$ while $g_n1_{X\setminus N}\to f1_{X\setminus N}$. We now define
	\[G_n\coloneqq g_n^{-1}(B\setminus\{0\}),\]
	which is in $\mc S$ has finite measure by \autoref{rem:better-simp-int}. In particular, $f(x)\ne0$ implies that either $x\in N$ or $g_n(x)\to f(x)$ as $n\to\infty$, which requires $g_n(x)\ne0$ for some $n$ and thus $x\in G_n$ for some $n$. As such, we see
	\[F\subseteq N\cup\bigcup_{n=1}^\infty G_n,\]
	which competes the proof of the claim.

	Now, $f(x)\ge1_E(x)$ almost everywhere, so select some $N'\in\mc S$ such that $\mu(N')=0$ and $x\in X\setminus N'$ implies $f(x)\ge1_E(x)$. With this in mind, we define
	\[E_n\coloneqq (E\setminus N')\cap\left(N\cup\bigcup_{i=1}^n G_i\right).\]
	In particular, we see that \autoref{lem:fin-additive-is-monotone} and \autoref{lem:finitely-additive-is-subaddtive} imply
	\[\mu(E_n)\le\mu(N')+\mu(N)+\sum_{i=1}^n\mu(G_i)\]
	is a finite sum of finite real numbers and is therefore finite.

	As such, we note $x\in E\setminus N'$ implies $f(x)\ne0$ and thus $x\in F$, so $E\setminus N'\subseteq F$, so $E\setminus N'=\bigcup_{n=1}^\infty E_n$. Further, we see $E_n\subseteq E_n\cup G_{n+1}=E_{n+1}$ straight from the definition, so \autoref{prop:measure-union-up} tells us
	\[\mu(E\setminus N')=\lim_{n\to\infty}\mu(E_n).\]
	However, $E_n\subseteq E\setminus N'$ implies $1_{E_n}(x)\le1_{E\setminus N'}(x)$ for each $x\in X$, so $1_{E_n}(x)\le f(x)$ for $x\in X\setminus N'$, so $1_{E_n}(x)\le f(x)$ almost everywhere, so \autoref{lem:bound-ints} tells us
	\[\int_X1_{E_n}\,d\mu\le\int_Xf\,d\mu.\]
	Noting $\mu(E_n)=\int_X1_{E_n}\,d\mu$ by \autoref{ex:integrate-indicator}, we see $\mu(E_n)\le\int_Xf\,d\mu$ for each $n$. It follows that
	\[\mu(E\setminus N')\le\int_Xf\,d\mu.\]
	However, $\mu(N')=0$, so $\mu(E\cap N')=0$ by \autoref{lem:fin-additive-is-monotone}, so $\mu(E\setminus N')=\mu(E)-\mu(E\cap N')=\mu(E)$. This finishes.
\end{proof}
\begin{corollary} \label{cor:bounded-simp-meas-is-simp-int}
	Fix a measure space $(X,\mc S,\mu)$ and a normed vector space $B$. Further, fix a simple $\mc S$-measurable function $f\colon X\to B$ and a $\mu$-integrable function $g\colon X\to\RR$. If $\norm{f(x)}\le g(x)$ almost everywhere, then $f$ is simple $\mu$-integrable.
\end{corollary}
\begin{proof}
	Fixing any $y\in(\im f)\setminus\{0\}$, we have to show that $f^{-1}(\{y\})$ has finite measure. Well, by \autoref{lem:fin-additive-is-monotone}, we can just show $E\coloneqq f^{-1}(B\setminus\{0\})$ has finite measure, where $E\in\mc S$ already.

	For this, we note that $\im f$ is finite, so $\{\norm{y}:y\in(\im f)\setminus\{0\}\}$ is finite and therefore has a minimum value $r$. Note $r>0$ because $\norm y=0$ implies $y=0$. As such, we note that
	\[r1_E(x)\le\norm{f}(x)\]
	for all $x\in X$ because either $x\notin E$ and thus $f(x)=0$ or $x\in E$ and thus $r\le\norm{f(x)}$. It follows $1_E(x)\le\frac1r\norm f(x)\le\frac1rg(x)$ almost everywhere, so \autoref{lem:bound-meas-set} tells us that $E$ has finite measure. In particular, $\frac 1rg$ is $\mu$-integrable by \autoref{lem:int-is-vec-space}. 
\end{proof}

\subsection{Convergence in Mean, Again}
We now move towards showing that $\mc L^1$ is complete. To state the result, we need to (re)define converging in mean for our $\mu$-integrable functions.
\begin{definition}[Converge in mean]
	Fix a measure space $(X,\mc S,\mu)$ and a Banach space $B$. Then a sequence $\{f_n\}_{n\in\NN}$ of $\mu$-integrable functions \textit{converges in mean} to a $\mu$-integrable function $f\colon X\to B$ if and only if $\norm{f-f_n}_1\to0$ as $n\to\infty$.
\end{definition}
\begin{definition}[Mean Cauchy]
	Fix a measure space $(X,\mc S,\mu)$ and a Banach space $B$. Then a sequence $\{f_n\}_{n\in\NN}$ of $\mu$-integrable functions is \textit{mean Cauchy} if and only if $\norm{f_m-f_n}_1\to0$ as $m,n\to\infty$.
\end{definition}
\begin{remark}
	If everything is simple $\mu$-integrable, then we note the fact that $\norm f_1$ is the same for $\mu$-integrable functions as for simple $\mu$-integrable functions means that our definitions above also do not change.
\end{remark}
\begin{remark} \label{rem:converge-in-mean-computes-int}
	Roughly speaking, convergence in mean lets us compute integrals. Namely, if $f_n\to f$ in mean, then we claim $\int_Xf_n\,d\mu\to\int_Xf\,d\mu$. Indeed, note
	\[\norm{\int_Xf\,d\mu-\int_Xf_n\,d\mu}=\norm{\int_X(f-f_n)\,d\mu}\le\int_X\norm{f-f_n}\,d\mu=\norm{f-f_n}_1\]
	by \autoref{prop:int-is-linear} and \autoref{lem:bound-norm-int}. Thus, for any $\varepsilon>0$, we use $f_n\to f$ in mean to find $N$ such that $n\ge N$ implies $\norm{f-f_n}_1<\varepsilon$, which implies $\norm{\int_Xf\,d\mu-\int_Xf_n\,d\mu}<\varepsilon$ as well. This finishes.
\end{remark}
We now take a deep breath and run a few checks. Here is a comparison result.
\begin{lemma} \label{lem:converge-in-mean-to-converge-in-measure}
	Fix a measure space $(X,\mc S,\mu)$ and a Banach space $(B,\norm\cdot)$. Further, fix a sequence of $\mu$-integrable functions $\{f_n\}_{n\in\NN}$ and another $\mu$-integrable function $f\colon X\to B$. If $f_n\to f$ in mean, then $f_n\to f$ in measure.
\end{lemma}
\begin{proof}
	We imitate \autoref{lem:mean-cauchy-is-cauchy-in-measure}; note the statement makes sense because the $f_n$ and $f$ are $\mc S$-measurable. Now, fix some $\varepsilon>0$. Then, for any $n$, we define
	\[E_{n}\coloneqq\{x\in X:\norm{f(x)-f_n(x)}\ge\varepsilon\}\]
	so that we want to show $\mu(E_n)\to0$ as $n\to\infty$; note $E_n$ is $\mc S$-measurable because all the relevant functions are $\mc S$-measurable. Well, we see that each $x\in X$ has
	\[1_{E_{n}}(x)\le\frac{\norm{f(x)-f_n(x)}}\varepsilon,\]
	so \autoref{lem:bound-meas-set} tells us that
	\[\mu(E_n)\le\int_X\frac{\norm{f(x)-f_n(x)}}\varepsilon\,d\mu=\frac1\varepsilon\int_X\norm{f(x)-f_n(x)}\,d\mu=\frac{\norm{f-f_n}_1}\varepsilon,\]
	where we have used \autoref{prop:int-is-linear}. Thus,
	\[\lim_{n\to\infty}\mu(E_n)\le\lim_{n\to\infty}\frac{\norm{f-f_n}_1}\varepsilon=\frac1\varepsilon\lim_{n\to\infty}\norm{f-f_n}_1,\]
	which is $0$ because $f_n\to f$ in mean. The fact that $\mu(E_n)\ge0$ for each $n$ tells us $\lim_{n\to\infty}\mu(E_n)\ge0$, so $\mu(E_n)\to0$ as $n\to\infty$ follows.
\end{proof}
Here is a nice consequence.
\begin{lemma} \label{lem:norm-zero-to-zero}
	Fix a measure space $(X,\mc S,\mu)$ and a normed vector space $B$. Given a $\mu$-integrable function $f\colon X\to B$, if $\norm f_1=0$, then $f(x)=0$ almost everywhere.
\end{lemma}
\begin{proof}
	Let $z$ denote the zero function so that we want to show $f(x)=0=z(x)$ almost everywhere. Note that the sequence $\{f_n\}_{n\in\NN}$ of functions defined by $f_n\coloneqq z$ are all simple $\mu$-integrable functions (vacuously). As such, we see that $f_n\to z$ in measure.
	
	On the other hand, we see $f_n\to f$ in mean because any $\varepsilon>0$ can set $N=0$ so that $n\ge N$ has
	\[\norm{f-f_n}_1=\norm{f-0}_1=\norm f_1=0<\varepsilon.\]
	However, $f_n\to f$ in mean implies that $f_n\to f$ in measure by \autoref{lem:converge-in-mean-to-converge-in-measure}. It follows from \autoref{lem:uniq-limit-in-measure} that $f(x)=z(x)$ almost everywhere.
\end{proof}
And here is the converse.
\begin{lemma} \label{lem:zero-to-norm-zero}
	Fix a measure space $(X,\mc S,\mu)$ and a Banach space $B$. Given some $\mu$-integrable function $f\colon X\to B$, if $f(x)=0$ almost everywhere, then $\norm f_1=0$.
\end{lemma}
\begin{proof}
	Define $f_n\colon X\to B$ to be the zero function for each $n$. The main claim is that $f_n\to f$ in measure. Indeed, because $f(x)=0$ almost everywhere, we can find $E\in\mc S$ such that $\mu(E)=0$ while $f(x)=0$ for $x\in X\setminus E$. As such, for any $\varepsilon>0$, we note that any $n\ge1$ has
	\[\{x:\norm{f(x)-f_n(x)}\ge\varepsilon\}=\{x:\norm{f(x)}\ge\varepsilon\}\subseteq\{x:f(x)\ne0\}\subseteq E.\]
	Thus, \autoref{lem:fin-additive-is-monotone} tells us $\mu\left(\{x:\norm{f(x)-f_n(x)}\ge\varepsilon\}\right)\le\mu(E)=0$, finishing.
	
	We now note that each $\norm{f_n}$ is the zero function and hence (vacuously) a simple $\mc S$-integrable function, and $\norm{f_n}\to\norm f$ by \autoref{lem:linear-combo-in-measure}. Thus, by definition of our integral,
	\[\norm f_1=\int_X\norm f\,d\mu=\lim_{n\to\infty}\int_X\norm{f_n}\,d\mu=\lim_{n\to\infty}\norm{f_n}_1=\lim_{n\to\infty}0=0.\]
	This is what we wanted.
\end{proof}
And here is the total result.
\begin{lemma} \label{lem:norm-diff-zero-iff-equal-ae}
	Fix a measure space $(X,\mc S,\mu)$ and a normed vector space $B$. Then two $\mu$-integrable functions $f,g\colon X\to B$ have $\norm{f-g}_1=0$ if and only if $f(x)=g(x)$ almost everywhere.
\end{lemma}
\begin{proof}
	We let $[h]\in L^1(X,\mc S,\mu,B)$ denote the equivalence class of a $\mu$-integrable function $h\colon X\to B$.

	In one direction, if $[f]=[g]$, then $f-g\in\mc N(X,\mc S,\mu,B)$, so $\norm{f-g}_1=0$. It follows that $f(x)-g(x)=0$ almost everywhere by \autoref{lem:norm-zero-to-zero}, so we can select $E\in\mc S$ such that $\mu(E)=0$ while $f(x)-g(x)=0$ for $x\in X\setminus E$. As such, $f(x)=g(x)$ for $x\in X\setminus E$, so $f(x)=g(x)$ almost everywhere.

	In the other direction, suppose $f(x)=g(x)$ almost everywhere. Then we can select $E\in\mc S$ such that $\mu(E)=0$ while $f(x)=g(x)$ for $x\in X\setminus E$. It follows $f(x)-g(x)=0$ for $x\in X\setminus E$, so $(f-g)(x)=0$ almost everywhere. Thus, $\norm{f-g}_1=0$ by \autoref{lem:zero-to-norm-zero}.
\end{proof}
\begin{remark} \label{rem:equal-ae-gives-equal-ints}
	Fix a measure space $(X,\mc S,\mu)$ and a normed vector space $B$. \autoref{lem:norm-diff-zero-iff-equal-ae} tells us that $\mu$-integrable functions $f,g\colon X\to B$ equal almost everywhere have $\norm{f-g}_1=0$. As an application, we note \autoref{prop:int-is-linear} and \autoref{lem:bound-norm-int} imply
	\[\norm{\int_Xf\,d\mu-\int_Xg\,d\mu}=\norm{\int_X(f-g)\,d\mu}\le\int_X\norm{f-g}\,d\mu=\norm{f-g}_1=0,\]
	so $\int_Xf\,d\mu=\int_Xg\,d\mu$ follows.
\end{remark}

\subsection{Completeness of \texorpdfstring{$L^1$}{ L1}}
And now for our feature presentation.
\begin{proposition} \label{prop:l1-complete}
	Fix a measure space $(X,\mc S,\mu)$ and a Banach space $B$. Then a mean Cauchy sequence $\{f_n\}_{n\in\NN}$ of $\mu$-integrable functions converges in mean to some $\mu$-integrable function $f\colon X\to B$.
\end{proposition}
\begin{proof}
	For each $n$, \autoref{cor:simple-int-are-dense} grants some simple $\mu$-integrable function $g_n\colon X\to B$ with $\norm{f_n-g_n}_1<1/n$. We now proceed in steps.
	\begin{enumerate}
		\item We claim that $\{g_n\}_{n\in\NN}$ is mean Cauchy. Well, fix any $\varepsilon>0$. Then we can find some $N_f$ such that $m,n\ge N_f$ implies
		\[\norm{f_m-f_n}_1<\frac\varepsilon3.\]
		Thus, we define $N\coloneqq\max\{N_f,3/\varepsilon\}$ so that $m,n\ge N$ implies (by \autoref{cor:l1-seminorm}) that
		\[\norm{g_m-g_n}_1\le\norm{f_m-g_m}_1+\norm{f_m-f_n}_1+\norm{f_n-g_n}_1<\frac1m+\frac\varepsilon3+\frac1n\le\frac2N+\frac\varepsilon3\le\frac{2\varepsilon}3+\frac\varepsilon3=\varepsilon,\]
		which finishes.
		\item Next, we construct the limit function. Because $\{g_n\}_{n\in\NN}$ is mean Cauchy, it is Cauchy in measure by \autoref{lem:mean-cauchy-is-cauchy-in-measure}, so there is a uniformly Cauchy subsequence $\{g_{n_i}\}_{i\in\NN}$ by \autoref{thm:rw}. However, this subsequence $\{g_{n_i}\}_{i\in\NN}$ will then converge to some $\mc S$-measurable $g\colon X\to B$ almost uniformly by \autoref{lem:almost-uniform-cauchy-converges}. Note that $g$ is then $\mu$-integrable by definition.
		\item It remains to show that $f_n\to g$ in mean. Well, note $g_{n_i}\to g$ almost uniformly, so $g_{n_i}\to g$ in measure by \autoref{lem:almost-uniform-to-in-measure}, so $g_{n_i}\to g$ in mean by \autoref{lem:almost-simple-int-are-dense}. (This step is why it is important for the $g_\bullet$ to be simple $\mu$-integrable!) Finishing up, we fix any $\varepsilon>0$ and note that there is $N_g$ such that $i\ge N_g$ implies
		\[\norm{g_{n_i}-g}_1<\varepsilon/3.\]
		Further, $\{f_n\}_{n\in\NN}$ is mean Cauchy, so there is $N_f$ such that $m,n\ge N_f$ implies
		\[\norm{f_m-f_n}_1<\varepsilon/3.\]
		In total, we set $N\coloneqq\max\{N_f,N_g,3/\varepsilon\}$. Then $n\ge N$ implies $n\ge N_f$ and $n_n\ge n\ge N\ge N_g$, so (using \autoref{cor:l1-seminorm} some more)
		\[\norm{g-f_n}_1\le\norm{g-g_{n_n}}_1+\norm{g_{n_n}-f_{n_n}}_1+\norm{f_{n_n}-f_n}_1<\frac\varepsilon3+\frac1{n_n}+\frac\varepsilon3\le\frac\varepsilon3+\frac1{3/\varepsilon}+\frac\varepsilon3=\varepsilon,\]
		which is what we wanted.
		\qedhere
	\end{enumerate}
\end{proof}
\begin{corollary} \label{cor:limit-of-mu-ints}
	Fix a measure space $(X,\mc S,\mu)$ and a Banach space $(B,\norm\cdot)$. Given a mean Cauchy sequence $\{f_n\}_{n\in\NN}$ of $\mu$-integrable functions converging to some $\mc S$-measurable function $f\colon X\to B$ almost everywhere, we know $f$ is $\mu$-integrable, and $f_n\to f$ in mean.
\end{corollary}
\begin{proof}
	The main annoyance here is proving that $f$ is actually $\mu$-integrable. As such, we divide the proof into two steps.
	\begin{enumerate}
		\item Note \autoref{prop:l1-complete} promises some $\mu$-integrable function $f'\colon X\to B$ such that $f_n\to f'$ in mean. However, $f_n\to f'$ in mean implies that $f_n\to f'$ in measure by \autoref{lem:converge-in-mean-to-converge-in-measure}, so $f_n\to f'$ almost everywhere, so $f=f'$ almost everywhere.

		However, $f'$ is already $\mu$-integrable, so there exists some mean Cauchy sequence $\{g_n\}_{n\in\NN}$ of simple $\mu$-integrable functions such that $g_n\to f'$ almost everywhere. It follows that $g_n\to f$ almost everywhere as well, so $f$ is in fact $\mu$-integrable.

		\item Now, $f=f'$ almost everywhere implies that $\norm{f-f'}_1=0$ by \autoref{lem:norm-diff-zero-iff-equal-ae}. Thus, for any $\varepsilon>0$, we use $f_n\to f'$ in mean to find $N$ such that $n\ge N$ implies
		\[\norm{f'-f_n}<\varepsilon.\]
		However, we now see $\norm{f-f_n}_1\le\norm{f-f'}_1+\norm{f'-f_n}<\varepsilon$ by \autoref{cor:l1-seminorm}, finishing.
		\qedhere
	\end{enumerate}
\end{proof}
In order to actually state this as a completeness result, we need to turn the semi-norm $\norm\cdot_1$ into an actual norm.
\begin{notation}
	Fix a measure space $(X,\mc S,\mu)$ and a Banach space $B$. We set $\mc N(X,\mc S,\mu,B)\coloneqq\{f\in\mc L^1(X,\mc S,\mu,B):\norm f_1=0\}$ and
	\[L^1(X,\mc S,\mu,B)\coloneqq\mc L^1(X,\mc S,\mu,B)/\mc N(X,\mc S,\mu,B).\]
\end{notation}
\begin{remark}
	Given $\mu$-integrable functions $f,g\colon X\to B$, we claim that the equivalence classes $[f],[g]\in L^1(X,\mc S,\mu,B)$ are equal if and only if $f(x)=g(x)$ almost everywhere. Indeed, $[f]=[g]$ if and only if $f-g\in\mc N(X,\mc S,\mu,B)$, which is equivalent to $\norm{f-g}_1=0$. However, by \autoref{lem:norm-diff-zero-iff-equal-ae}, $\norm{f-g}_1=0$ is equivalent to $f(x)=g(x)$ almost everywhere.
\end{remark}
\begin{lemma}
	Fix a measure space $(X,\mc S,\mu)$ and a Banach space $B$. The function $\norm\cdot_1$ descends to a norm on $L^1(X,\mc S,\mu,B)$.
\end{lemma}
\begin{proof}
	This is a direct consequence \autoref{prop:semi-norm-to-norm}, applied to $\norm\cdot_1$ on $\mc L^1(X,\mc S,\mu,B)$.
\end{proof}
\begin{corollary} \label{cor:l1-complete}
	Fix a measure space $(X,\mc S,\mu)$ and a Banach space $B$. Then $L^1(X,\mc S,\mu,B)$ is the completion of the vector space of simple $\mu$-integrable functions.
\end{corollary}
\begin{proof}
	The normed vector space $L^1(X,\mc S,\mu,B)$ is complete by \autoref{prop:l1-complete}. Further, the space of simple $\mu$-integrable functions (modded out by the functions of norm zero) are dense in $L^1(X,\mc S,\mu,B)$ by \autoref{cor:simple-int-are-dense}.
\end{proof}
Next class we will begin trying to compute integrals.

\end{document}