% !TEX root = ../notes.tex

\documentclass[../notes.tex]{subfiles}

\begin{document}

\section{November 4}

We continue our journey towards integrating functions.

\subsection{Sequences Converging in Measure}
We pick up some basic tools on sequences converging in measure.
\begin{lemma} \label{lem:converge-in-measure-uniq-ae}
	Fix a normed vector space $(B,\norm\cdot)$ and a measure space $(X,\mc S,\mu)$. Now, suppose a sequence $\{f_n\}_{n\in\NN}$ of $\mc S$-measurable functions converges to both $f$ and $g$ in measure, where $f$ and $g$ are both $\mc S$-measurable. Then $f=g$ almost everywhere; i.e., $\{x\in X:f(x)\ne g(x)\}$ is a null set.
\end{lemma}
\begin{proof}
	Before we do anything at all, we note that $f-g$ is $\mc S$-measurable by \autoref{lem:meas-is-vec-space}, so
	\[N\coloneqq\{x\in X:f(x)\ne g(x)\}=(f-g)^{-1}(B\setminus\{0\})\]
	is $\mc S$-measurable by \autoref{cor:meas-has-meas-pre-image}.
	
	Now, fix any $\varepsilon>0$; we show $\mu(N)<\varepsilon$. The key observation is that
	\[\norm{f(x)-g(x)}\le\norm{f(x)-f_n(x)}+\norm{f_n(x)-g(x)},\]
	so it follows that $\norm{f(x)-g(x)}\ge\varepsilon$ forces $\norm{f(x)-f_n(x)}\ge\varepsilon/2$ or $\norm{g(x)-g_n(x)}\ge\varepsilon/2$. Thus,
	\[\{x:\norm{f(x)-g(x)}\ge\varepsilon\}\subseteq\{x:\norm{f(x)-f_n(x)}\ge\varepsilon/2\}\cup\{x:\norm{g(x)-f_n(x)}\ge\varepsilon/2\},\]
	so \autoref{lem:finitely-additive-is-subaddtive} tells us
	\[\mu(\{x\in X:\norm{f(x)-g(x)}\ge\varepsilon\})\le\mu(\{x\in X:\norm{f(x)-f_n(x)}\ge\varepsilon/2\})+\mu(\{x\in X:\norm{g(x)-f_n(x)}\ge\varepsilon/2\}).\]
	But now, as $n\to\infty$, we see that the right-hand side goes to $0+0=0$ because $f_n\to f$ and $f_n\to g$ in measure, so it follows that
	\begin{equation}
		\mu(\{x\in X:\norm{f(x)-g(x)}\ge\varepsilon\})=0. \label{eq:norm-big-vanishes-equal-limits}
	\end{equation}
	We now send $\varepsilon\to0^+$. Namely, we see $f(x)\ne g(x)$ is equivalent to $\norm{f(x)-g(x)}>0$ is equivalent to $\norm{f(x)-g(x)}\ge1/n$ for some $n\in\NN$, so
	\[N\coloneqq\{x\in X:f(x)\ne g(x)\}=\{x\in X:\norm{f(x)-g(x)}>0\}=\bigcup_{n\in\NN}\{x\in X:\norm{f(x)-g(x)}\ge1/n\}.\]
	Thus,
	\[\mu(N)\le\sum_{n\in\NN}\mu(\{x\in X:\norm{f(x)-g(x)}\ge1/n\})\stackrel*=\sum_{x\in X}0=0,\]
	so $N$ is in fact a null set. Notably, $\stackrel*=$ has used \autoref{eq:norm-big-vanishes-equal-limits}.
\end{proof}
\begin{lemma} \label{lem:linear-combo-in-measure}
	Fix a normed $k$-vector space $(B,\norm\cdot)$ and a measure space $(X,\mc S,\mu)$. Fix sequences of $\mc S$-measurable functions $\{f_n\}_{n\in\NN}$ and $\{g_n\}_{n\in\NN}$ with $f_n\to f$ and $g_n\to g$ in measure as $n\to\infty$.
	\begin{listalph}
		\item We have $f_n+g_n\to f+g$ in measure.
		\item Given some scalar $a\in k$, we have $af_n\to af$ in measure.
		\item We have $\norm{f_n}\to\norm f$ as $n\to\infty$.
	\end{listalph}
\end{lemma}
\begin{proof}
	We go ahead and let $|\cdot|$ denote the norm on base field $k$ of $B$.
	\begin{listalph}
		\item Note that the $f_n+g_n$ and $f+g$ are all $\mc S$-measurable by \autoref{lem:meas-is-vec-space}.
		
		Now, by the triangle inequality, we see
		\[\norm{(f(x)+g(x))-(f_n(x)+g_n(x))}\le\norm{f(x)-f_n(x)}+\norm{g(x)-g_n(x)}.\]
		We now proceed as in \autoref{lem:converge-in-measure-uniq-ae}. Fix $\varepsilon>0$. If the left-hand side exceeds $\varepsilon$, then one of the terms on the right-hand side must exceed $\varepsilon/2$, so
		\begin{align*}
			\{x:\norm{(f(x)+g(x))-(f_n(x)+g_n(x))}\ge\varepsilon\}&\subseteq\{x:\norm{f(x)-f_n(x)}\ge\varepsilon/2\} \\
			&\qquad\cup\{x:\norm{g(x)-g_n(x)}\ge\varepsilon/2\}.
		\end{align*}
		Thus, \autoref{lem:finitely-additive-is-subaddtive} tells us
		\begin{align*}
			\mu(\{x:\norm{(f(x)+g(x))-(f_n(x)+g_n(x))}\ge\varepsilon\}) &\le \mu(\{x:\norm{f(x)-f_n(x)}\ge\varepsilon/2\}) \\
			&\qquad+\mu(\{x:\norm{g(x)-g_n(x)}\ge\varepsilon/2\}).
		\end{align*}
		However, $\varepsilon/2>0$, so taking $n\to\infty$ and using our convergence in measure tells us that
		\[\lim_{n\to\infty}\mu(\{x:\norm{(f(x)+g(x))-(f_n(x)+g_n(x))}\ge\varepsilon\})\le0+0=0,\]
		so we are done after noting that $\mu$ will only output nonnegative values, so the limit is at least nonnegative.

		\item Note that the $af_n$ and $af$ are all $\mc S$-measurable by \autoref{lem:meas-is-vec-space}.
		
		Now, fix some $\varepsilon>0$ so that we want to show that
		\[L\coloneqq\lim_{n\to\infty}\mu(\{x\in X:\norm{af_n(x)-af(x)}\ge\varepsilon\})\stackrel?=0.\]
		If $a=0$, then $af_n(x)=af(x)=0$ for all $x\in X$, so $\{x\in X:\norm{af_n(x)-af(x)}\ge\varepsilon\}$ is empty, so the result follows.

		Otherwise, take $a\ne0$ so that $|a|>0$. Now, note $\norm{af_n(x)-af(x)}=|a|\cdot\norm{f_n(x)-f(x)}$, so it follows $\norm{af_n(x)-af(x)}\ge\varepsilon$ if and only if $\norm{f_n(x)-f(x)}\ge\varepsilon/|a|$. Thus,
		\[L=\lim_{n\to\infty}\mu(\{x\in X:\norm{f_n(x)-f(x)}\ge\varepsilon/|a|\}).\]
		However, $\varepsilon/|a|>0$ because $\varepsilon>0$, so the above limit vanishes because $f_n\to f$ in measure as $n\to\infty$.

		\item Observe that the $g_n$ and $g$ are all $\mc S$-measurable by \autoref{cor:take-norms-is-measurable}.

		Now, fix some $\varepsilon>0$. By the (reverse) triangle inequality,
		\[|\norm{f(x)}-\norm{f_n(x)}|\le\norm{f(x)-f_n(x)},\]
		so any $\varepsilon>0$ has
		\[\{x:|\norm{f(x)}-\norm{f_n(x)}|\ge\varepsilon\}\subseteq\{x:\norm{f(x)-f_n(x)}\ge\varepsilon\}.\]
		Thus, \autoref{lem:fin-additive-is-monotone} tells us
		\[\lim_{n\to\infty}\mu(\{x:|\norm{f(x)}-\norm{f_n(x)}|\ge\varepsilon\})\le\lim_{n\to\infty}\mu(\{x:\norm{f(x)-f_n(x)}\ge\varepsilon\}).\]
		The right-hand limit vanishes because $f_n\to f$ in measure, so the left-hand limit must vanish as well because the limit's terms are nonnegative.
		\qedhere
	\end{listalph}
\end{proof}
Here is the analogous result for sequences Cauchy in measure.
\begin{lemma} \label{lem:linear-combo-cauchy-in-measure}
	Fix a normed $k$-vector space $(B,\norm\cdot)$ and a measure space $(X,\mc S,\mu)$. Fix sequences of $\mc S$-measurable functions $\{f_n\}_{n\in\NN}$ and $\{g_n\}_{n\in\NN}$ which are Cauchy in measure.
	\begin{listalph}
		\item The sequence $\{f_n+g_n\}_{n\in\NN}$ is Cauchy in measure.
		\item Given some scalar $a\in k$, the sequence $\{af_n\}_{n\in\NN}$ is Cauchy in measure.
		\item The sequence of functions $\{\norm{f_n}\}_{n\in\NN}$ is Cauchy in measure.
	\end{listalph}
\end{lemma}
\begin{proof}
	These proofs are essentially the same as \autoref{lem:linear-combo-in-measure} with the appropriate names changed. Again, we let $|\cdot|$ denote the norm on base field $k$ of $B$.
	\begin{listalph}
		\item Note that the $f_n+g_n$ are all $\mc S$-measurable by \autoref{lem:meas-is-vec-space}.
		
		Now, by the triangle inequality, we see
		\[\norm{(f_m(x)+g_m(x))-(f_n(x)+g_n(x))}\le\norm{f_m(x)-f_n(x)}+\norm{g_m(x)-g_n(x)}.\]
		Fix $\varepsilon>0$. As usual
		\begin{align*}
			\{x:\norm{(f_m(x)+g_m(x))-(f_n(x)+g_n(x))}\ge\varepsilon\}&\subseteq\{x:\norm{f_m(x)-f_n(x)}\ge\varepsilon/2\} \\
			&\qquad\cup\{x:\norm{g_m(x)-g_n(x)}\ge\varepsilon/2\},
		\end{align*}
		so \autoref{lem:finitely-additive-is-subaddtive} tells us
		\begin{align*}
			\mu(\{x:\norm{(f_m(x)+g_m(x))-(f_n(x)+g_n(x))}\ge\varepsilon\}) &\le \mu(\{x:\norm{f_m(x)-f_n(x)}\ge\varepsilon/2\}) \\
			&\qquad+\mu(\{x:\norm{g_m(x)-g_n(x)}\ge\varepsilon/2\}).
		\end{align*}
		However, $\varepsilon/2>0$, so taking $m,n\to\infty$ and using our Cauchy in measure conditions tells us that
		\[\lim_{m,n\to\infty}\mu(\{x:\norm{(f_m(x)+g_m(x))-(f_n(x)+g_n(x))}\ge\varepsilon\})\le0+0=0,\]
		so we are done after noting that $\mu$ will only output nonnegative values, so the limit is at least nonnegative.

		\item Note that the $af_n$ are all $\mc S$-measurable by \autoref{lem:meas-is-vec-space}.
		
		Now, fix some $\varepsilon>0$ so that we want to show that
		\[L\coloneqq\lim_{m,n\to\infty}\mu(\{x\in X:\norm{af_m(x)-af_n(x)}\ge\varepsilon\})\stackrel?=0.\]
		If $a=0$, then $af_n(x)=af(x)=0$ for all $x\in X$, so $\{x\in X:\norm{af_m(x)-af_n(x)}\ge\varepsilon\}$ is empty, so the result follows.

		Otherwise, take $a\ne0$ so that $|a|>0$. Now, note $\norm{af_m(x)-af_n(x)}=|a|\cdot\norm{f_m(x)-f_n(x)}$, so it follows $\norm{af_m(x)-af_n(x)}\ge\varepsilon$ if and only if $\norm{f_m(x)-f_n(x)}\ge\varepsilon/|a|$. Thus,
		\[L=\lim_{m,n\to\infty}\mu(\{x\in X:\norm{f_m(x)-f_n(x)}\ge\varepsilon/|a|\}).\]
		However, $\varepsilon/|a|>0$ because $\varepsilon>0$, so the above limit vanishes because $\{f_n\}_{n\in\NN}$ is Cauchy in measure.

		\item Observe that the $g_n$ are all $\mc S$-measurable by \autoref{cor:take-norms-is-measurable}.

		Now, fix some $\varepsilon>0$. By the (reverse) triangle inequality,
		\[|\norm{f_m(x)}-\norm{f_n(x)}|\le\norm{f_m(x)-f_n(x)},\]
		so any $\varepsilon>0$ has
		\[\{x:|\norm{f_m(x)}-\norm{f_n(x)}|\ge\varepsilon\}\subseteq\{x:\norm{f_m(x)-f_n(x)}\ge\varepsilon\}.\]
		Thus, \autoref{lem:fin-additive-is-monotone} tells us
		\[\lim_{m,n\to\infty}\mu(\{x:|\norm{f_m(x)}-\norm{f_n(x)}|\ge\varepsilon\})\le\lim_{m,n\to\infty}\mu(\{x:\norm{f_m(x)-f_n(x)}\ge\varepsilon\}).\]
		The right-hand limit vanishes because $\{f_n\}_{n\in\NN}$ is Cauchy in measure, so the left-hand limit must vanish as well because the limit's terms are also nonnegative.
		\qedhere
	\end{listalph}
\end{proof}
We will want a few results on subsequences later on.
\begin{lemma} \label{lem:in-measure-subsequence}
	Fix a normed vector space $(B,\norm\cdot)$ and a measure space $(X,\mc S,\mu)$. Now, fix$\mc S$-measurable functions $\{f_n\}_{n\in\NN}$ and $f$. If $f_n\to f$ in measure as $n\to\infty$, then $f_{n_i}\to f$ in measure as $i\to\infty$ for any subsequence $\{f_{n_i}\}_{i\in\NN}$.
\end{lemma}
\begin{proof}
	Fix some $\varepsilon>0$. Then any $\delta>0$ has some $N$ for which $n\ge N$ has
	\[\mu(\{x\in X:\norm{f(x)-f_n(x)}\ge\varepsilon\})<\delta.\]
	As such, for any $i\ge N$, we see $n_i\ge i\ge N$, so
	\[\mu(\{x\in X:\norm{f(x)-f_{n_i}(x)}\ge\varepsilon\})<\delta,\]
	which finishes.
\end{proof}
\begin{lemma} \label{lem:subsequence-has-same-limit-in-measure}
	Fix a measure space $(X,\mc S,\mu)$ and a normed vector space $B$. Further, fix a sequence of simple $\mc S$-measurable functions $\{f_n\}_{n\in\NN}$ which is Cauchy in measure. If a subsequence $\{f_{n_i}\}_{i\in\NN}$ converges to a function $f$ in measure, then the full sequences $\{f_n\}_{n\in\NN}$ converges to $f$ in measure.
\end{lemma}
\begin{proof}
	Fix any $\varepsilon>0$ and $\delta>0$. We need $N$ such that $n\ge N$ implies
	\[0\le\mu(\{x\in X:\norm{f(x)-f_n(x)}\ge\varepsilon\})\stackrel?<\delta.\]
	Well, we note that any $n$ and $i$ will have
	\[\norm{f(x)-f_n(x)}\le\norm{f(x)-f_{n_i}(x)}+\norm{f_{n_i}(x)-f_n(x)},\]
	so
	\[\{x\in X:\norm{f(x)-f_n(x)}\ge\varepsilon\}\subseteq\{x\in X:\norm{f(x)-f_{n_i}(x)}\ge\varepsilon/2\}\cup\{x\in X:\norm{f_{n_i}(x)-f_n(x)}\ge\varepsilon/2\}.\]
	Now, $\{f_n\}_{n\in\NN}$ being Cauchy in measure allows us to pick $N$ such that $m,n\ge N$ implies
	\[\mu(\{x\in X:\norm{f_m(x)-f_n(x)}\ge\varepsilon/2\})<\frac\delta2.\]
	Additionally, $f_{n_i}\to f$ in measure grants $N'$ such that $i\ge N'$ implies
	\[\mu(\{x\in X:\norm{f(x)-f_{n_i}(x)}\ge\varepsilon/2\})<\frac\delta2.\]
	Thus, for any $n\ge N$, we select any $i\ge\max\{N,N'\}$. Notably, $n_i\ge i\ge N$ as well, so
	\[\mu(\{x\in X:\norm{f(x)-f_n(x)}\ge\varepsilon\})<\frac\delta2+\frac\delta2=\delta,\]
	where we have used the above inequalities in addition to \autoref{lem:finitely-additive-is-subaddtive}.
\end{proof}
Lastly, here is the expected uniqueness results.
\begin{lemma} \label{lem:uniq-limit-in-measure}
	Fix a normed vector space $(B,\norm\cdot)$ and a measure space $(X,\mc S,\mu)$. Further, fix a sequence $\{f_n\}_{n\in\NN}$ of $\mc S$-measurable functions converging in measure to an $\mathcal S$-measurable function $f\colon X\to B$. Given an $\mathcal S$-measurable function $g\colon X\to B$, we have $f=g$ almost everywhere if and only if $f_n\to g$ in measure.
\end{lemma}
\begin{proof}
	In one direction, suppose $f=g$ almost everywhere so that we have some $E\in\mc S$ with $\mu(E)=0$ such that $f(x)\ne g(x)$ implies $x\in E$. Now, fix some $\varepsilon>0$. For any $\delta>0$, we are promised $N$ such that $n\ge N$ implies
	\[\mu(\{x\in X:\norm{f(x)-f_n(x)}\ge\varepsilon\})<\delta.\]
	Now, we note that $\norm{g(x)-f_n(x)}\ge\varepsilon$ implies that either $f(x)\ne g(x)$ so that $x\in E$ or $\norm{f(x)-f_n(x)}\ge\varepsilon$, so it follows
	\[\{x\in X:\norm{g(x)-f_n(x)}\ge\varepsilon\}=\{x\in X:\norm{f(x)-f_n(x)}\ge\varepsilon\}\cup\{x\in E:\norm{g(x)-f_n(x)}\ge\varepsilon\}.\]
	Notably, $\mu(\{x\in E:\norm{g(x)-f_n(x)}\ge\varepsilon\})=0$ as a subset of $E$, so we see that
	\[\mu(\{x\in X:\norm{g(x)-f_n(x)})\le\{x\in X:\norm{f(x)-f_n(x)}\ge\varepsilon\}+0<\delta,\]
	which finishes this direction.

	Conversely, suppose $f_n\to g$ in measure. Note that $f(x)\ne g(x)$ if and only if $\norm{f(x)-g(x)}>0$ if and only if $\norm{f(x)-g(x)}\ge1/m$ for some positive integer $m$. Thus,
	\[\{x\in X:f(x)\ne g(x)\}\subseteq\bigcup_{m=1}^\infty\{x\in X:\norm{f(x)-g(x)}\ge1/m\}.\]
	By \autoref{lem:premeas-is-countable-subadd}, to show that $\{x\in X:f(x)\ne g(x)\}$ is a null set, it suffices to show that $\{x\in X:\norm{f(x)-g(x)}\ge1/m\}$ is a null set. Well, note that any positive integer $n$ has
	\[\norm{f(x)-g(x)}\le\norm{f(x)-f_n(x)}+\norm{g(x)-f_n(x)},\]
	implying
	\[\{x:\norm{f(x)-g(x)}\ge1/m\}\subseteq\{x:\norm{f(x)-f_n(x)}\ge1/(2m)\}\cup\{x:\norm{g(x)-f_n(x)}\ge1/(2m)\}.\]
	Now, for any $\delta>0$, because $f_n\to f$ and $f_n\to g$ in measure, we are promised some $N$ large enough so that $n\ge N$ has
	\[\mu(\{x:\norm{f(x)-f_n(x)}\ge1/(2m)\}),\mu(\{x:\norm{g(x)-f_n(x)}\ge1/(2m)\})<\frac\delta2.\]
	It follows by \autoref{lem:finitely-additive-is-subaddtive} that
	\[\mu(\{x\in X:\norm{f(x)-g(x)}\ge1/m\})<\frac\delta2+\frac\delta2=\delta\]
	for any $\delta>0$. Thus, $\mu(\{x\in X:\norm{f(x)-g(x)}\ge1/m\})=0$ follows.
\end{proof}

\subsection{Restricting Measurable Functions}
Analogously to \autoref{lem:linear-combo-in-measure} and \autoref{lem:linear-combo-cauchy-in-measure}, we have the following.
\begin{lemma} \label{lem:restrict-converge-in-measure}
	Fix a normed vector space $(B,\norm\cdot)$ and a measure space $(X,\mc S,\mu)$, and fix some $E\in\mc S$. Given a sequence $\{f_n\}_{n\in\NN}$ of $\mc S$-measurable functions with $f_n\to f$ in measure as $n\to\infty$, then $f_n1_E\to f1_E$ in measure as $n\to\infty$.
\end{lemma}
\begin{proof}
	Note that the $f_n1_E$ and $f1_E$ are all $\mc S$-measurable by \autoref{lem:restrict-meas-functions}, so the claim at least makes sense.

	For brevity, we set $g_n\coloneqq\norm{f1_E-f_n1_E}$ for each $n$. We would like to show
	\[\lim_{n\to\infty}\mu\left(g_n^{-1}([\varepsilon,\infty))\right)\stackrel?=0.\]
	If $x\notin E$, then note $g_n(x)=0$; otherwise, $g_n(x)=\norm{f(x)-f_n(x)}$ because $1_E(x)=1$. As such, for $\varepsilon>0$, we see $x\in g_n^{-1}([\varepsilon,\infty))$ requires $x\in E$ and then $\norm{f(x)-f_n(x)}\ge\varepsilon$; conversely, $x\in E$ with $\norm{f(x)-f_n(x)}\ge\varepsilon$ does give $g_n(x)\ge\varepsilon$.

	Thus, we note that
	\[g_n^{-1}([\varepsilon,\infty))\subseteq\{x:\norm{f(x)-f_n(x)}\ge\varepsilon\},\]
	so \autoref{lem:fin-additive-is-monotone} tells us
	\[\lim_{n\to\infty}\mu\left(g_n^{-1}([\varepsilon,\infty))\right)\le\lim_{n\to\infty}\mu\left(\{x:\norm{f(x)-f_n(x)}\ge\varepsilon\}\right),\]
	where the right-hand limit vanishes because $f_n\to f$ in measure as $n\to\infty$. Thus, the left-hand limit also vanishes because the terms of the limit are nonnegative.
\end{proof}
\begin{lemma}
	Fix a normed vector space $(B,\norm\cdot)$ and a measure space $(X,\mc S,\mu)$, and fix some $E\in\mc S$. Given a sequence $\{f_n\}_{n\in\NN}$ of $\mc S$-measurable functions which is Cauchy in measure, then $\{f_n1_E\}_{n\in\NN}$ is still Cauchy in measure.
\end{lemma}
\begin{proof}
	As usual, the proof is exactly the same as before. Note that the $f_n1_E$ and $f1_E$ are all $\mc S$-measurable by \autoref{lem:restrict-meas-functions}, so the claim at least makes sense.

	For brevity, we set $g_{m,n}\coloneqq\norm{f_m1_E-f_n1_E}$ for each $m$ and $n$. We would like to show
	\[\lim_{m,n\to\infty}\mu\left(g_{m,n}^{-1}([\varepsilon,\infty))\right)\stackrel?=0.\]
	If $x\notin E$, then note $g_{m,n}(x)=0$; otherwise, $g_{m,n}(x)=\norm{f_m(x)-f_n(x)}$ because $1_E(x)=1$. As such, for $\varepsilon>0$, we see $x\in g_{m,n}^{-1}([\varepsilon,\infty))$ requires $x\in E$ and then $\norm{f_m(x)-f_n(x)}\ge\varepsilon$; conversely, $x\in E$ with $\norm{f_m(x)-f_n(x)}\ge\varepsilon$ does give $g_{m,n}(x)\ge\varepsilon$.

	Thus, we note that
	\[g_{m,n}^{-1}([\varepsilon,\infty))\subseteq\{x:\norm{f_m(x)-f_n(x)}\ge\varepsilon\},\]
	so \autoref{lem:fin-additive-is-monotone} tells us
	\[\lim_{m,n\to\infty}\mu\left(g_{m,n}^{-1}([\varepsilon,\infty))\right)\le\lim_{m,n\to\infty}\mu\left(\{x:\norm{f_m(x)-f_n(x)}\ge\varepsilon\}\right),\]
	where the right-hand limit vanishes because $\{f_{n}\}_{n\in\NN}$ is Cauchy in measure. Thus, the left-hand limit also vanishes because the terms of the limit are nonnegative.
\end{proof}
The above corollary promises the following notation.
\begin{notation}
	Fix a normed vector space $(B,\norm\cdot)$ and a measure space $(X,\mc S,\mu)$. Then a simple integrable function $f$ on $X$ and $E\in\mc S$ will have
	\[\int_Ef\,d\mu\coloneqq\int_Xf1_E\,d\mu.\]
\end{notation}
\begin{remark}
	One can define
	\[\mu_f(E)\coloneqq\int_Ef\,d\mu,\]
	and it is not too hard to check that this defines a measure on $\mc S$ which is valued in $B$. This $\mu_f$ will later be called the ``indefinite integral for $f$.'' We will postpone writing this out until we are ready to talk about what this looks like when $f$ is a general $\mu$-integrable function instead of a simple $\mu$-integrable function.
\end{remark}

\subsection{Almost Uniform Convergence}
As we tend to do, we now return to a context which is perhaps too general.
\begin{definition}[Almost uniformly]
	Fix a normed vector space $B$ and a measure space $(X,\mc S,\mu)$. Then a sequence of functions $f_n\colon X\to B$ for $n\in\NN$ converges \textit{almost uniformly} to $f$ if and only if every $\varepsilon>0$ has some $E^\varepsilon\in\mc S$ such that $\mu(E^\varepsilon)<\varepsilon$ and $f_n|_{X\setminus E}\to f|_{X\setminus E}$ uniformly.
\end{definition}
\begin{remark}
	The term ``almost'' above is different from the ``almost everywhere'' that we've been seeing.
\end{remark}
As usual, with a convergence definition, we have a Cauchy definition.
\begin{definition}[Almost uniformly Cauchy]
	Fix a measure space $(X,\mc S,\mu)$ and a normed vector space $B$. Then a sequence of functions $f_n\colon X\to B$ for $n\in\NN$ is \textit{almost uniformly Cauchy} if and only if every $\varepsilon>0$ has some $E^\varepsilon\in\mc S$ such that $\mu(E^\varepsilon)<\varepsilon$ and $\{f_n|_{X\setminus E}\}_{n\in\NN}$ is uniformly Cauchy.
\end{definition}
We take a deep breath and run some of our usual checks.
\begin{lemma} \label{lem:uniq-almost-uniform-limit}
	Fix a normed vector space $B$ and a measure space $(X,\mc S,\mu)$. Now, suppose a sequence $\{f_n\}_{n\in\NN}$ converges almost uniformly to a function $f$. Then, for a function $g\colon X\to B$, we have $f=g$ almost everywhere if and only if $f_n\to g$ almost uniformly.
\end{lemma}
\begin{proof}
	In one direction, suppose $f=g$ almost everywhere so that $\{x\in X:f(x)\ne g(x)\}$ is contained in some $N\in\mc S$ such that $\mu(N)=0$. Then for any $\varepsilon>0$, we note that $f_n\to f$ almost uniformly promises $F\in\mc S$ such that $\mu(X\setminus F)<\varepsilon$ while $f_n|_F\to f|_F$ uniformly. Now, $f|_{X\setminus N}=g|_{X\setminus N}$, so we note $f_n|_{F\setminus N}\to g|_{F\setminus N}$ uniformly (by restricting $f_n|_F\to f|_F$ uniformly) while
	\[\mu(X\setminus(F\setminus N))=\mu\big((X\setminus F)\cup N\big)=\mu(X\setminus F)+\mu(N\cap F),\]
	where $\mu(N\cap F)=0$ because $\mu(N)=0$.

	The other direction is harder. Define $N\coloneqq\{x\in X:f(x)\ne g(x)\}$, and we show that $N$ is a null set. Well, for any $d>0$, we are promised subsets $F_d,G_d\in\mc S$ such that $\mu(F_d),\mu(G_d)<1/d$ and $f_n|_{X\setminus F}\to f|_{X\setminus F}$ and $f_n|_{X\setminus G}\to g|_{X\setminus G}$ uniformly as $n\to\infty$.

	In particular, if $x\notin(F_d\cup G_d)$, then our uniform convergence will imply pointwise convergence at $x$, so $f_n(x)\to f(x)$ and $f_n(x)\to g(x)$ as $n\to\infty$. It follows that $f(x)=g(x)$ by properties of convergence. Apply contraposition, we conclude that $N\subseteq F_d\cup G_d$; as such, we use \autoref{lem:finitely-additive-is-subaddtive} to note that
	\[\mu(F_d\cup G_d)\le\mu(F_d)+\mu(G_d)<\frac2d.\]
	We now send $d\to\infty$. Define
	\[E\coloneqq\bigcap_{d\ge1}(F_d\cup G_d),\]
	which lives in $\mc S$ because $\mc S$ is a $\sigma$-ring. As above, we see that $N\subseteq F_d\cup G_d$ for each $d$, so $N\subseteq E$. Further, $E\subseteq F_d\cup G_d$ tells us by \autoref{lem:fin-additive-is-monotone} that
	\[\mu(E)\le\mu(F_d\cup G_d)<\frac2d\]
	for any positive integer $d$. In particular, sending $d\to\infty$ forces $\mu(E)=0$, which finishes the proof that $N$ is a null set.
\end{proof}
\begin{lemma} \label{lem:almost-uniformly-subsequence}
	Fix a normed vector space $B$ and a measure space $(X,\mc S,\mu)$. Given a sequence of functions $\{f_n\}_{n\in\NN}$ converging to $f\colon X\to B$ almost uniformly, we have $f_{n_i}\to f$ almost uniformly for any subsequence $\{f_{n_i}\}_{i\in\NN}$.
\end{lemma}
\begin{proof}
	For any $\varepsilon>0$, we are promised $E\in\mc S$ such that $\mu(X\setminus E)<\varepsilon$ and $f_n\to f$ uniformly on $X\setminus E$. This means that any $\delta>0$ has some $N$ such that $n\ge N$ implies
	\[|f_n(x)-f(x)|<\delta\]
	for each $x\in X\setminus E$. However, this implies that $i\ge N$ gives $n_i\ge i\ge N$ and thus
	\[|f_n(x)-f(x)|<\delta\]
	for each $x\in X\setminus E$, so $f_{n_i}\to f$ uniformly on $X\setminus E$ as well.
\end{proof}
\begin{lemma} \label{lem:almost-uniform-linear-combo}
	Fix a normed $k$-vector space $(B,\norm\cdot)$ and a measure space $(X,\mc S,\mu)$. Fix sequences of functions $f_n\colon X\to B$ and $g_n\colon X\to B$ with $f_n\to f$ and $g_n\to g$ almost uniformly as $n\to\infty$.
	\begin{listalph}
		\item We have $f_n+g_n\to f+g$ almost uniformly.
		\item Given some scalar $a\in k$, we have $af_n\to af$ almost uniformly.
		\item We have $\norm{f_n}\to\norm f$ almost uniformly as $n\to\infty$.
	\end{listalph}
\end{lemma}
\begin{proof}
	We go ahead and let $|\cdot|$ denote the norm on $k$. For any $\varepsilon>0$, we will also go ahead and let $F_\varepsilon,G_\varepsilon\in\mc S$ denote the subsets of $X$ with $\mu(F_\varepsilon),\mu(G_\varepsilon)<\varepsilon$ for which $f_n|_{X\setminus F_\varepsilon}\to f$ and $g_n|_{X\setminus G_\varepsilon}\to g$ uniformly.
	\begin{listalph}
		\item For any $\varepsilon>0$, define $E_\varepsilon\coloneqq F_{\varepsilon/2}\cup G_{\varepsilon/2}$. Namely, $E_\varepsilon\in\mc S$, and by \autoref{lem:finitely-additive-is-subaddtive}, we see
		\[\mu(E_\varepsilon)\le\mu(F_{\varepsilon/2})+\mu(G_{\varepsilon/2})=\varepsilon.\]
		As such, we claim that $(f_n+g_n)|_{X\setminus E_\varepsilon}\to(f+g)|_{X\setminus E_\varepsilon}$ uniformly as $n\to\infty$.

		Well, for any $\delta>0$, we are promised $N_f$ such that any $n\ge N_f$ and $x\notin F_{\varepsilon/2}$ will have
		\[\norm{f(x)-f_n(x)}<\delta/2.\]
		We are promised an analogous constant $N_g$ for $g_n$ going to $g$, so we set $N\coloneqq\max\{N_f,N_g\}$. Then $n\ge N$ implies $n\ge N_f$ and $n\ge N_g$; as such, if $x\notin E_\varepsilon$, then $x\notin F_{\varepsilon/2}$ and $x\notin G_{\varepsilon/2}$, so
		\[\norm{(f+g)(x)-(f_n+g_n)(x)}\le\norm{f(x)-f_n(x)}+\norm{g(x)-g_n(x)}<\frac\delta2+\frac\delta2=\delta.\]

		\item If $a=0$, then we see that $af_n=af=0$. As such, for any $\varepsilon>0$, we set $E=\emp$ so that $\mu(E)=0$ while $af_n\to af$ uniformly as $n\to\infty$ because $af_n(x)=0=af(x)$ for any $x\in X$.

		Otherwise, we have $a\ne0$ and so $|a|>0$. Thus, for any $\varepsilon>0$, we note $F_\varepsilon$ will have $\mu(F_\varepsilon)<\varepsilon$, so we claim that $(af_n)|_{X\setminus F_\varepsilon}\to(af)|_{X\setminus F_\varepsilon}$ uniformly as $n\to\infty$.

		Well, we already know that $f_n|_{X\setminus F_\varepsilon}\to f|_{X\setminus F_\varepsilon}$ uniformly as $n\to\infty$. Thus, for any $\delta>0$, there is a constant $N$ so that any $n\ge N$ and $x\notin F_\varepsilon$ will have
		\[\norm{f(x)-f_n(x)}<\frac\delta{|a|}.\]
		It follows that $n\ge N$ and $x\notin F_\varepsilon$ gives
		\[\norm{(af)(x)-(af_n)(x)}=|a|\cdot\norm{f(x)-f_n(x)}<|a|\cdot\frac\delta{|a|}=\delta.\]

		\item Unsurprisingly, for any $\varepsilon>0$, we note that $F_\varepsilon$ has $\mu(F_\varepsilon)<\varepsilon$, so we claim that $\norm f_n|_{X\setminus F_\varepsilon}\to\norm f|_{X\setminus F_\varepsilon}$ almost uniformly as $n\to\infty$.

		Well, we know that $f_n|_{X\setminus F_\varepsilon}\to f|_{X\setminus F_\varepsilon}$ as $n\to\infty$. Thus, for any $\delta>0$, we are promised a constant $N$ such that $n\ge N$ and $x\notin F_\varepsilon$ will have
		\[\norm{f(x)-f_n(x)}<\delta.\]
		As such, we note that the (reverse) triangle inequality gives
		\[|\norm{f(x)}-\norm{f_n(x)}|\le\norm{f(x)-f_n(x)}<\delta,\]
		which finishes.
		\qedhere
	\end{listalph}
\end{proof}
\begin{lemma}
	Fix a normed $k$-vector space $(B,\norm\cdot)$ and a measure space $(X,\mc S,\mu)$. Fix sequences of functions $f_n\colon X\to B$ and $g_n\colon X\to B$ which are almost uniformly Cauchy.
	\begin{listalph}
		\item The sequence $\{f_n+g_n\}_{n\in\NN}$ is almost uniformly Cauchy.
		\item Given some scalar $a\in k$, the sequence $\{af_n\}_{n\in\NN}$ is almost uniformly Cauchy.
		\item The sequence $\{\norm{f_n}\}_{n\in\NN}$ is almost uniformly Cauchy.
	\end{listalph}
\end{lemma}
\begin{proof}
	As usual, these proofs are basically the same.
	
	We go ahead and let $|\cdot|$ denote the norm on $k$. For any $\varepsilon>0$, we will also go ahead and let $F_\varepsilon,G_\varepsilon\in\mc S$ denote the subsets of $X$ with $\mu(F_\varepsilon),\mu(G_\varepsilon)<\varepsilon$ for which $\{f_n|_{X\setminus F_\varepsilon}\}_{n\in\NN}$ and $\{g_n|_{X\setminus G_\varepsilon}\}_{n\in\NN}$ are uniformly Cauchy.
	\begin{listalph}
		\item For any $\varepsilon>0$, define $E_\varepsilon\coloneqq F_{\varepsilon/2}\cup G_{\varepsilon/2}$. Namely, $E_\varepsilon\in\mc S$, and by \autoref{lem:finitely-additive-is-subaddtive}, we see
		\[\mu(E_\varepsilon)\le\mu(F_{\varepsilon/2})+\mu(G_{\varepsilon/2})=\varepsilon.\]
		As such, we claim that $\{(f_n+g_n)|_{X\setminus E_\varepsilon}\}_{n\in\NN}$ is uniformly Cauchy.

		Well, for any $\delta>0$, we are promised $N_f$ such that any $m,n\ge N_f$ and $x\notin F_{\varepsilon/2}$ will have
		\[\norm{f_m(x)-f_n(x)}<\delta/2.\]
		We are promised an analogous constant $N_g$ for $g_n$ going to $g$, so we set $N\coloneqq\max\{N_f,N_g\}$. Then $m,n\ge N$ implies $m,n\ge N_f$ and $m,n\ge N_g$; as such, if $x\notin E_\varepsilon$, then $x\notin F_{\varepsilon/2}$ and $x\notin G_{\varepsilon/2}$, so
		\[\norm{(f_m+g_m)(x)-(f_n+g_n)(x)}\le\norm{f_m(x)-f_n(x)}+\norm{g_m(x)-g_n(x)}<\frac\delta2+\frac\delta2=\delta.\]

		\item If $a=0$, then we see that $af_n=af_m=0$. As such, for any $\varepsilon>0$, we set $E=\emp$ so that $\mu(E)=0$ while $\{af_n\}_{n\in\NN}$ is uniformly Cauchy because $af_n(x)=0=af_m(x)$ for any $x\in X$.

		Otherwise, we have $a\ne0$ and so $|a|>0$. Thus, for any $\varepsilon>0$, we note $F_\varepsilon$ will have $\mu(F_\varepsilon)<\varepsilon$, so we claim that $\{(af_n)|_{X\setminus F_\varepsilon}\}_{n\in\NN}$ is uniformly Cauchy.

		Well, we already know that $\{f_n|_{X\setminus F_\varepsilon}\}_{n\in\NN}$ is uniformly Cauchy. Thus, for any $\delta>0$, there is a constant $N$ so that any $m,n\ge N$ and $x\notin F_\varepsilon$ will have
		\[\norm{f_m(x)-f_n(x)}<\frac\delta{|a|}.\]
		It follows that $m,n\ge N$ and $x\notin F_\varepsilon$ gives
		\[\norm{(af_m)(x)-(af_n)(x)}=|a|\cdot\norm{f_m(x)-f_n(x)}<|a|\cdot\frac\delta{|a|}=\delta.\]

		\item Unsurprisingly, for any $\varepsilon>0$, we note that $F_\varepsilon$ has $\mu(F_\varepsilon)<\varepsilon$, so we claim that $\{\norm f_n|_{X\setminus F_\varepsilon}\}_{n\in\NN}$ is uniformly Cauchy.

		Well, we know that $\{f_n|_{X\setminus F_\varepsilon}\}_{n\in\NN}$ is uniformly Cauchy. Thus, for any $\delta>0$, we are promised a constant $N$ such that $m,n\ge N$ and $x\notin F_\varepsilon$ will have
		\[\norm{f_m(x)-f_n(x)}<\delta.\]
		As such, we note that the (reverse) triangle inequality gives
		\[|\norm{f_m(x)}-\norm{f_n(x)}|\le\norm{f_m(x)-f_n(x)}<\delta,\]
		which finishes.
		\qedhere
	\end{listalph}
\end{proof}
Now, here is the main result, which we will not prove today.
\begin{restatable}[Riesz--Weyl]{thm}{rwthm} \label{thm:rw}
	Fix a measure space $(X,\mc S,\mu)$ and a normed vector space $B$. Let $\{f_n\}_{n\in\NN}$ be a sequence of $\mc S$-measurable $B$-valued functions which are Cauchy in measure. Then there is a subsequence $\{f_n\}_{n\in\NN}$ which is almost uniformly Cauchy.
\end{restatable}
\noindent In particular, we will be able to define a limit function for the sequence $\{f_n\}_{n\in\NN}$ outside some null set, which will finally allow us to take limits of simple integrable functions in a way that makes sense.

\end{document}