% !TEX root = ../notes.tex

\documentclass[../notes.tex]{subfiles}

\begin{document}

\section{March 10}

Today we talk about cohomology of projective space.

\subsection{Cohomology of Projective Space}
Fix a Noetherian ring $A$, and we work with $\PP^r_A=\Proj A[x_0,\ldots,x_r]$. Here is our result; note that we are looking at our cohomology as $A$-modules, which we know is legal because we can take flasque resolutions and value everything in $\mathrm{Mod}_A$ instead of merely $\mathrm{Ab}$.
\begin{theorem} \label{thm:proj-cohom}
	Fix a Noetherian ring $A$, and set $X\coloneqq\PP^r_A$ with $r>0$. Further, set $S\coloneqq A[x_0,\ldots,x_n]$ and $\mc F\coloneqq\bigoplus_{r\in\ZZ}\OO(n)$.
	\begin{listalph}
		\item We have $H^0(X,\mc F)=S$. In other words, $H^0(X,\OO(n))=S_n$ for any $n$.
		\item For any $0<i<r$, we have $H^i(X,\mc F)=0$.
		\item We have $H^r(X,\OO(-r-1))\cong A$.
		\item For all $n$, there is a perfect $A$-pairing
		\[H^0(X,\OO(n))\otimes_AH^r(X,\OO(-n-r-1))\to H^r(X,\OO(-r-1))=A.\]
		In other words, $H^0(X,\OO(n))\cong\op{Hom}_A(H^r(X,\OO(-n-r-1)),A)$.
	\end{listalph}
\end{theorem}
\begin{remark}
	Because $\PP^r_A$ can be covered by $r+1$ affine open subschemes in the standard affine open cover $\mf U$, we see
	\[H^i(X,\mc F)=\check H^i(\mf U,\mc F)=0\]
	for $i>r$ by definition of \v Cech cohomology. Thus, we have no higher cohomology than given in \Cref{thm:proj-cohom}.
\end{remark}
\begin{remark}
	To properly state the perfect pairing, we note that a global section $s\in H^0(X,\OO(n))$ produces a map $s\colon\OO(-n-r-1)\to\OO(-r-1)$ by multiplication and checking degrees, which by functoriality produces a map $s\colon H^r(X,\OO(-n-r-1))\to H^r(X,\OO(-r-1))$. The perfect pairing is asserting that the induced map
	\[H^0(X,\OO(n))\cong\op{Hom}_A(H^r(X,\OO(-n-r-1)),A)\]
	is an isomorphism. Notably, by adjunction, we are really talking about a map $s\otimes t\mapsto st$, which is what is included in the statement.
\end{remark}
\begin{proof}[Proof of \Cref{thm:proj-cohom}]
	We showed (a) last class in \Cref{prop:h0-proj-space}. Now, we use \v Cech cohomology. Set $U_i\coloneqq D_+(x_i)\cong\Spec A[x_{0/i},\ldots,x_{n/i}]/(x_{i/i}-1)$ so that $\mf U\coloneqq\{U_i\}$ is an open cover of $X$. Notably, we see that
	\[U_{i_0\cdots i_p}=D_+(x_{i_0}\cdots x_{i_p}).\]
	Computing our sections now, we see
	\[\Gamma(U_{i_0\cdots i_p},\mc F)=\bigoplus_{n\in\ZZ}\Gamma(U_{i_0\cdots i_p},\OO(n))=\bigoplus){n\in\ZZ}\Gamma(D_+(x_{i_0}\cdots x_{i_p}),\OO(n)).\]
	Now, accounting for our shifting, $\Gamma(D_+(x_{i_0}\cdots x_{i_p}),\OO(n))$ is asking for degree-$0$ elements of $(S_n)_{x_{i_0}\cdots x_{i_p}}$, which are really the degree-$n$ elements of $S_{x_{i_0}\cdots x_{i_p}}$. As such,
	\[\Gamma(U_{i_0\cdots i_p},\mc F)=\bigoplus_{n\in\ZZ}(S_{x_{i_0}\cdots x_{i_p}})_n=S_{x_{i_0}\cdots x_{i_p}}.\]
	As such, our \v Cech complex $C^\bullet(\mf U,\mc F)$ looks like
	\[0\to\prod_{i=0}^rS_{x_i}\xrightarrow{d^0}\prod_{0<i_0<i_1\le r}S_{x_{i_0}x_{i_1}}\xrightarrow{d^1}\cdots\to\prod_{0<i_0<\cdots<i_r\le r}S_{x_{i_0}\cdots x_{i_r}}\xrightarrow{d^{r-1}}S_{x_0\cdots x_r}\to0.\]
	For example, $\im d^{r-1}$ is the span of the monomials $x_0^{\nu_1}\cdots x_r^{\nu_r}$ where at least one of the $\nu_\bullet$ is nonnegative because it must arise from some $S_{x_0\cdots\widehat{x_k}\cdots x_r}$, which means that $x_k$ is not used. As such, we can compute
	\[\check H^r(\mf U,\mc F)=\frac{S_{x_0\cdots x_r}}{\im d^{r-1}}.\]
	In particular, the numerator is a free $A$-module generated by monomials $x_0^{\nu_1}\cdots x_r^{\nu_r}$, but the denominator is a free $A$-module generated by monomials with at least one positive coordinate. Thus, by taking the quotient appropriately, $\check H^r(\mf U,\mc F)$ is a free $A$-module generated by $x_0^{\nu_0}\cdots x_r^{\nu_r}$ where all the terms are (strictly) negative!

	For example, if we insist on looking at $\check H^r(\mf U,\OO(-r-1))$, then we are looking at an $A$-module freely generated by $x_0^{\nu_0}\cdots x_r^{\nu_r}$ where each $\nu_\bullet$ is negative, and $\nu_0+\cdots+\nu_r=-r-1$. However, the only way to do this is for everything to be $-1$, so we see
	\[H^r(X,\OO(-r-1))=\check H^r(\mf U,\OO(-r-1))=Ax_0^{-1}\cdots x_r^{-1}.\]
	This is (c).

	One can see (d) similarly. In particular, $H^0(X,\OO(n))$ has an $A$-basis of monomials $x_0^{\nu_0'}\cdots x_r^{\nu_r'}$ with $\nu_\bullet'\ge0$ and total degree $n$; on the other hand, $H^r(X,\OO(-n-r-1))$ has an $A$-basis of monomials $x_0^{\nu_0}\cdots x_r^{\nu_r}$ with $\nu_\bullet\le0$ and total degree $-n-r-1$, so we can simply multiply these homogeneous polynomials to produce a homogeneous polynomial of degree $-r$, which we can put into $H^r(X,\OO(-r-1))$. In other words, we are sending
	\[x_0^{\nu_0'}\cdots x_r^{\nu_r'}\cdot x_0^{\nu_0}\cdots x_r^{\nu_r}\mapsto x_0^{\nu_0'+\nu_0}\cdots x_r^{\nu_r'+\nu_r}\pmod{\im d^{r-1}}.\]
	It remains to show that we are producing a perfect pairing in this way. In other words, given some morphism $H^r(X,\OO(-n-r-1))\to H^r(X,\OO(-r-1))$, it must arise from some unique element of $S$ via the above multiplication; it suffices to check this on our basis of monomials. One can see this from a direct computation: given some monomial $x_0^{\nu_0}\cdots x_r^{\nu_r}$ satisfying
	\[\sum_{i=0}^r(-\ell_i-1)=n,\]
	we need to produce a monomial $x_0^{\nu_0'}\cdots x_{\nu_r'}$ such that the corresponding products land us in degree $-r-1$. However, we see that we must have $0\le\nu_i\le-\nu_i-1$ everywhere to ensure that we land in a monomial with negative degrees everywhere (up to some residue class---anything with a nonnegative degree somewhere will vanish entirely), but then equalities must hold everywhere, which makes our chosen monomial unique.

	Lastly, we must show (b). Localizing our \v Cech complex at $x_r$ to compute the cohomology of $\mc F|_{D_+(x_r)}$ via the standard affine open covering using essentially the same \v Cech complex. (There are some redundant terms that already had $x_r$ in our tuples, but they do not matter.) However, $\mc F|_{D_+(x_r)}$ is a quasicoherent sheaf on the affine scheme $D_+(x_r)$, so its cohomology at index $0<i<r$ vanishes, so we see
	\[H^i(X,\mc F)=\check H^i(\mf U,\mc F)=H^i(U_r,\mc F)=0.\]
	Next time we will show that $x_r\colon H^i(U_r)\to H^i(U_r,\mc F)$ is an isomorphism, which will complete the computation.
\end{proof}

\end{document}