% !TEX root = ../notes.tex

\documentclass[../notes.tex]{subfiles}

\begin{document}

\section{February 13}

We hopefully finish discussing $H^2(\op{Gal}(L/K),L^\times)$ for local fields $L$ today.

\subsection{Finishing \texorpdfstring{$H^2$}{ H2}}
To finish up the computation of $H^2$, we make a final remark.
\begin{proposition}
	Fix a local field $K$, and let $I\subseteq\op{Gal}(K^{\mathrm{sep}}/K)$ be the kernel of the restriction map $\op{Gal}(K^{\mathrm{sep}}/K)\to\op{Gal}(K^{\mathrm{unr}}/K)$. Then $H^2(I,K^{\mathrm{sep}\times})=0$.
\end{proposition}
\begin{proof}
	We claim that the image of $H^2(I,\OO_{K^{\mathrm{sep}}}^\times)\to H^2(I,K^{\mathrm{sep}\times})$ vanishes. Indeed, fix some class $[f]$ in there, and because we have defined our cohomology as a colimit, it can be exhibited as from some finite extension as
	\[[f_L]\in H^2(\op{Gal}(L/K^{\mathrm{unr}}),\OO_L^\times).\]
	We can now show that $[f_L]$ is in the image of the map $H^1(\op{Gal}(L/K^{\mathrm{unr}}),\op{PGL}_n(\OO_L))\to H^2(\op{Gal}(L/K^{\mathrm{unr}}),L^\times)$ for some $n$, which will be enough by the usual exact sequence. Indeed, we have a formula: set $n\coloneqq[L:K^{\mathrm{unr}}]$ and $G\coloneqq\op{Gal}(L/K^{\mathrm{unr}})$, and let $f_L$ be the desired cocycle. Then our proof of the vanishing of $H^2(\op{Gal}(L/K^{\mathrm{unr}}),\OO_L^\times)$ tells us that we are in fact in the image of $H^1(\op{Gal}(L/K^{\mathrm{unr}}),\op{PGL}_n(\OO_L))$; explicitly, we have
	\[M\coloneqq\bigoplus_{s\in G}\OO_Le_s,\]
	where the $G$-action is given by $\rho_g\colon e_s\mapsto f_L(g,s)$. This $M$ is providing an element of $H^1(\op{Gal}(L/K^{\mathrm{unr}}),\op{PGL}_n(\OO_L))$, which produces\footnote{It turns out that we can just take this as $\PP_{\OO_L}M$.} an $\mathcal O_{K^{\mathrm{unr}}}$-scheme $P$ which is isomorphic to $\PP^{n-1}_{\OO_L}$ over $\OO_L$. But then last class we showed that this scheme must have vanishing class, so we are done.

	Continuing, we actually claim that the map $H^2(I,\OO_{K^{\mathrm{sep}}}^\times)\to H^2(I,K^{\mathrm{sep}\times})$ is surjective. Indeed, we have some exact sequence
	\[0\to\OO_{K^{\mathrm{sep}}}\to K^{\mathrm{sep}\times}\stackrel v\to\QQ\to0,\]
	where $v$ is our valuation. Now, $H^2(I,\QQ)$ vanishes because $\QQ$ is a divisible group (indeed, take any cocycle, using the colimit forces it into a finite extension, and then divisibility causes the cocycle to have vanishing class there). As such, our exact sequence will do the trick here. This finishes the proof.
\end{proof}
\begin{remark}
	Let us explain further where $H^i(G,\QQ)=0$ for any finite group $G$ and $i>0$. Indeed, the point is that the forgetful functor
	\[(-)^G\colon\mathrm{Mod}_{\QQ[G]}\to\mathrm{Vec}_\QQ\]
	is exact, which is enough because it causes our cohomology to vanish. Well, we note that we have two additive functors
	\[M^G\into M\xrightarrow{\frac1{\#G}\sum_{g\in G}g}M^G\]
	whose composite is the identity. In particular, using our lift, we get a canonical decomposition $M\cong M'\otimes M^G$, which tells us that $(-)^G$ should be an exact functor.
\end{remark}

\subsection{Back to Global Things}
Fix an extension of global fields $L/K$ with Galois group $G$. Given a place $v$ of $K$, we note that we have a decomposition
\[L\otimes_KK_v\cong\prod_{w\mid v}L_w,\]
where $L_w$ is the completion of $L$ at some place $w$ over $v$.
\begin{proposition}
	Fix everything as above. For fixed $L$-place $w_0\mid v$, we have a $G$-module isomorphism
	\[L\otimes_KK_v\cong\op{Ind}_{G_{w_0}}^GL_{w_0}.\]
\end{proposition}
\begin{proof}
	Recall $\op{Ind}_{G_{w_0}}^GL_{w_0}=\op{Mor}_{G_{w_0}}(G,L_{w_0})$. As such, we choose representatives $g_1,\ldots,g_r$ for $G/G_{w_0}$ to see that
	\[\op{Ind}_{G_{w_0}}^GL_{w_0}\cong\prod_{i=1}^rL_{w_0}g_i,\]
	where the $i$th component dictates where $g_i$ goes. But now the point is that the right-hand side remains $r$ different copies of the completion $L_{w_0}$, which is in fact the same as $L\otimes_KK_v$ above. One should check that this commutes with the $G$-action, but indeed it does.
\end{proof}
\begin{remark}
	One can replace everything with units as $L_w^\times$.
\end{remark}
\begin{corollary} \label{cor:get-to-induced}
	Fix everything as above. For fixed $L$-place $w_0\mid v$, we have a canonical isomorphism
	\[H^\bullet\Bigg(G,\prod_{w\mid v}L_w^\times\Bigg)\cong H^\bullet(G_{w_0},L_{w_0}^\times).\]
\end{corollary}
\begin{proof}
	More generally, given a subgroup $H\subseteq G$, we have the sequence of functors
	\[\mathrm{Mod}_H\xrightarrow{\op{Ind}}\mathrm{Mod}_G\xrightarrow{(-)^G}\mathrm{Ab}.\]
	Now, the functor $\mathrm{Ind}$ is exact and takes injectives to injectives (it has an exact left adjoint given by the restriction functor), so we can compute cohomology for given $H$-module either directly or by inducing first. The result follows.
\end{proof}

\end{document}