% !TEX root = ../notes.tex

\documentclass[../notes.tex]{subfiles}

\begin{document}

\section{April 17}

We continue discussing the Bombieri--Vinogradov theorem.

\subsection{The Bombieri--Vinogradov Theorem}
Recall the following statement.
\begin{theorem}
	For all $A>0$ and $Q\in\left[\sqrt x(\log x)^{-A},x^{1/2}\right]$, we have
	\[\sum_{q\le Q}\max_{y\le x}\max_{\substack{1\le a\le q\\\gcd(a,q)=1}}\left|\psi(y;q,a)-\frac y{\varphi(q)}\right|\ll\sqrt xQ(\log x)^5.\]
\end{theorem}
\begin{proof}
	During the Friday lecture, we found
	\[\max_{\substack{1\le a\le q\\\gcd(a,q)=1}}\left|\psi(y;q,a)-\frac y{\varphi(q)}\right|\le\frac1{\varphi(q)}\sum_\chi|\psi'(y,\chi)|,\]
	for some function $\psi'$. We would like to only use primitive characters, so observe that if $\chi$ is an imprimitive character induced by $\overline\chi$, then we can upper-bound the difference as
	\[|\psi'(y,\chi)-\psi'(y,\overline\chi)|\ll(\log qy)^2,\]
	so we achieve
	\[S\coloneqq\max_{\substack{1\le a\le q\\\gcd(a,q)=1}}\left|\psi(y;q,a)-\frac y{\varphi(q)}\right|\ll(\log qy)^2+\frac1{\varphi(q)}\sum_\chi|\psi'(y,\overline\chi)|.\]
	Summing gives
	\[\sum_{q\le Q}\max_{y\le x}\max_{\substack{1\le a\le q\\\gcd(a,q)=1}}\left|\psi(y;q,a)-\frac y{\varphi(q)}\right|\ll\sum_{q\le Q} Q(\log qx)^2+{\sum_{\chi\pmod q}}^*\max_{y\le x}|\psi'(y,\chi)|\sum_{kq\le Q}\frac1{\varphi(kq)}.\]
	We now note that $\varphi(kq)\ge\varphi(k)\varphi(q)$, so the last summation can be controlled like a harmonic series. Namely, we observe
	\[\sum_{k\le z}\frac1{\varphi(k)}\le\prod_{p\le z}\sum_{\nu=0}^\infty\frac1{\varphi(p^\nu)}=\prod_{p\le z}\left(1+\frac1{p-1}\sum_{\nu=0}^\infty\frac1{p^\nu}\right)=\prod_{p\le z}\frac1{1-\frac1p}\left(1+\frac1{p(p-1)}\right)\ll\log z.\]
	Notably, the second term here converges absolutely, and the $\frac1{1-\frac1p}$ term produces the $\log z$. Inputting this inequality grants
	\[S\ll Q(\log Qx)^2+\log x\sum_{q\le Q}{\sum_\chi}^*\max_{y\le x}|\psi'(y,\chi)|^2.\]
	Thus, it remains to show
	\[\sum_{q\le Q}\frac1{\varphi(q)}{\sum_\chi}^*\max_{y\le x}|\psi'(y,\chi)|\ll\sqrt xQ(\log x)^4.\]
	Instead, we will prove the ``dyadic'' inequality
	\[\sum_{U\le q\le 2U}\frac1{\varphi(q)}{\sum_\chi}^*\max_{y\le x}|\psi'(y,\chi)|\ll\left(\frac xU+x^{5/6}+x^{1/2}U\right)(\log x)^4.\]
	To see how this implies the desired inequality, we sum over our dyadic intervals: note that the sum of the $x^{5/6}$ terms do not matter (they are strictly less than $x^{1/2}Q$). Similarly, the $x^{1/2}U$ term will sum to $\ll x^{1/2}Q$ and also does not matter. The last term is harder. To write it out, for $U\in[Q_1,Q]$ for some $Q_1$, we have
	\[\sum_{Q_1\le q\le Q}\frac1{\varphi(q)}{\sum_\chi}^*\max_{y\le x}|\psi'(y,\chi)|\ll(\log x)^4\left(\frac x{Q_1}+x^{5/6}\log x+x^{1/2}Q\right).\]
	For example, taking $Q_1=(\log x)^A$, we note that Siegel's theorem lets us bound
	\[|\psi'(y,\chi)|\ll x(\log x)^{-2A-4},\]
	so we can deal with the small values of $q$ as
	\[\sum_{q\le(\log x)^A}\frac1{\varphi(q)}{\sum_\chi}^*\max_{y\le x}|\psi'(y,\chi)|\ll x(\log x)^{-A-4},\]
	which is good enough for our purposes.

	It remains to prove our dyadic bound. We split this into two parts. To begin, we use the large sieve. We claim that
	\begin{align*}
		&\sum_{q\le Q}\frac q{\varphi(q)}{\sum_\chi}^*\max_U\bigg|\max_{\substack{1\le m\le M\\1\le n\le N\\mn\le U}}a_mb_n\chi(mn)\bigg| \ll \\
		&\qquad\left(M+Q^2\right)^{1/2}\left(N+Q^2\right)^{1/2}\log(2MN)\Bigg(\sum_{1\le m\le M}|a_m|^2\Bigg)^{1/2}\Bigg(\sum_{1\le n\le N}|b_n|^2\Bigg)^{1/2}.
	\end{align*}
	The main difficulty is dealing with the $mn\le U$ constraint. Indeed, without this requirement, we could use Cauchy--Schwarz by writing
	\begin{align*}
		\sum_{q\le Q}\frac q{\varphi(q)}{\sum_\chi}^*\Bigg|\sum_{1\le m\le M}\sum_{1\le n\le N}a_mb_n\chi(mn)\Bigg| &\le \Bigg(\sum_{q\le Q}\frac q{\varphi(q)}{\sum_\chi}^*\Bigg|\sum_{1\le m\le M}a_m\chi(m)\Bigg)^{1/2}\cdot \\
		&\qquad\qquad\Bigg(\sum_{q\le Q}\frac q{\varphi(q)}{\sum_\chi}^*\Bigg|\sum_{1\le n\le N}b_n\chi(n)\Bigg)^{1/2}.
	\end{align*}
	Then the large sieve grants us
	\[\ll\left(M+Q^2\right)^{1/2}\left(N+Q^2\right)^{1/2}\Bigg(\sum_{1\le m\le M}|a_m|^2\Bigg)^{1/2}\Bigg(\sum_{1\le n\le N}|b_n|^2\Bigg)^{1/2}.\]
	To achieve the desired inequality, the point is to ``complete the sum.'' Intuitively, we are essentially adding in a term of $1_{mn\le U}$ to our sum in order to bound via Fourier analysis. Namely, the Mellin transform of $1_{[0,1]}(t)$ is $\frac1s$. We will continue this next class.
\end{proof}

\end{document}